---
title: "HIV Cascades Calculations"
author: "Richard T. Gray"
date: Date - `r format(Sys.Date(), format="%B %d %Y")`
output: 
  word_document:
    reference_docx: mystyles.docx
---

This document describes and contains the code for producing the HIV care cascade estimates for Australia. It is written in dynamic format using R markdown v2 within Rstudio (version 0.98.1091 with R version`r substr(R.Version()$version.string,1,15)`) to produce a Word (.docx) document (see <http://rmarkdown.rstudio.com>). All the R code to clean the data and produce the figures and tables is contained in this file but suppressed in the final output document. 

```{r initialization, echo = FALSE, messages = FALSE, include=FALSE}
# Clear workspace
rm(list=ls()) 

# Libraries for data manipulation
require(readxl)
require(dplyr)
require(tidyr)
require(utils)

# Libraries for output
require(ggplot2)    
require(knitr)      
# require(gridExtra)  
require(RColorBrewer)

# Set working directory
setwd(file.path(path.expand("~"),"Research","!Evaluation_Modelling","project_care_cascades"))

# Base directory
basePath <- getwd()
dataFolder <- file.path(basePath,"data")

# Folder for plots 
outputFolder <- file.path(basePath,"output")

# Primary script parameters
currTime <- format(Sys.time(), "%y-%m-%d") # to append to files
analysisYear <- 2013
startYear <- 1990
# reclean <- FALSE  # Rerun data cleaning if true otherwise load already cleaned file.
pop <- "all"  # Population exposure category
region <- "all" # State for analysis

# TODO: Think about doing cumulative diagnoses for each state and population everytime

```

## Estimate cumulative HIV notifications 

```{r startAnalysis}
# Load clean file
hivdata <- read.csv(file.path(dataFolder, paste("hivnotifications",
                    toString(analysisYear),".csv", sep ="")))

# Setup overarching analysis                   
years <- startYear:analysisYear # Years to calculate and plot 

# Setup extra columns for analysis
hivdata$yeardiagnosis <- as.numeric(substr(hivdata$datediagnosis,1,4))
hivdata$yeardeath <- as.numeric(substr(hivdata$datedeath,1,4))

allyears <- min(hivdata$yeardiagnosis):analysisYear # All years of data

# Quick checks that we are not doing analysis outside of the data.
if (max(hivdata$yeardiagnosis) < analysisYear) {
  stop("No data for specified for final year of analysis.")
}

if (min(hivdata$yeardiagnosis) > startYear) {
  stop("No data for specified for first year of analysis.")
}

# Setup our working data frame
hivset <- filter(hivdata,yeardiagnosis <= analysisYear)

```

```{r plotcode, echo=FALSE,messages = FALSE,include=FALSE}
# Default plot specifications
graphics.off()

# Baseline theme for plot variables
opts <- theme(text = element_text(face = "bold",size=12,colour="black"),
  axis.text.x = element_text(face = "plain",size=10,colour="black"),
  axis.text.y = element_text(face = "plain",size=10,colour="black"),
  axis.line = element_line(colour="black"),
  axis.ticks = element_line(colour="black"),
  legend.position = "top",
  legend.background = element_rect(),
  legend.key = element_blank(),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"),
  plot.title=element_text(size=12, face="bold")
)

# Theme for pie chart
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  axis.text.y=element_blank(),
  plot.title=element_text(size=12, face="bold")
  )

# Setup colours
getPalette <- colorRampPalette(brewer.pal(12,"Set3"))
piecolours1 <- getPalette(12)
piecolours2 <- gray.colors(3)
piecolours <- c(piecolours2,piecolours1)
linecolours <- brewer.pal(4,"Set1")
```

Sort out diagnoses ...

```{r diagnoses}
# Setup dataframe for overall results by first extracting diagnoses details

# Overall annual results
hivresults <- hivset %>% 
  group_by(yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(totalnotifications  = cumsum(notifications)) %>% 
  filter(yeardiagnosis <= analysisYear)   

# Calculate the proportion of diagnoses by exposure and population category by 
# state and nationally. Theses data frames are used for further calculations

hivstate <- hivset %>% group_by(state,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(totalnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(proptotalnotifications = totalnotifications/sum(totalnotifications)) %>% 
  filter(yeardiagnosis <= analysisYear)   

hivexpnat <- hivset %>% group_by(expgroup,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(totalnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(proptotalnotifications = totalnotifications/sum(totalnotifications)) %>% 
  filter(yeardiagnosis <= analysisYear)   

hivexpstate <- hivset %>% group_by(expgroup,state,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(totalnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(state,yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(proptotalnotifications = totalnotifications/sum(totalnotifications)) %>% 
  filter(yeardiagnosis <= analysisYear)   

# TODO: Will have to do the same by population group
```

```{r functions}
# Load and define useful functions for analysis

# Load function to remove duplicates
source('~/Research/!Evaluation_Modelling/project_care_cascades/code/removeDuplicates.R', echo=TRUE)

# Load function to calculate number living with diagnosed HIV
source('~/Research/!Evaluation_Modelling/project_care_cascades/code/livingDiagnosed.R', echo=TRUE)

# Function to easily extract sub-populations of interest
subhivset <- function(hivdataframe, born = 'all', 
                      region = 'all', mode = 'all' ) {
  # Options
  # pop - all, seasia, ssafrica, indigenous, nonindigenous
  # state - nsw, vic, qld, wa, nt, tas, sa, act
  # expgroup - msm, pwid, hetero, other
  
  subframe <- hivdataframe
  #   if (born ~= 'all')
  #     subframe <- filter(subframe, state == born)
  if (region != 'all') {
    subframe <- filter(subframe, state == region)
  }

  if (mode != 'all') {
    subframe <- filter(subframe, expgroup == mode)            
  }

  # Return final subframe
  return(subframe)
}

# Function for removing duplicates annually
numUnique <- function(dobframe, years,ignore, file = NA) {
  # Loop through years and calculate cumulative number of unique cases 
  numcases <- rep(NA,length(years))
  for (ii in seq(along=years)) {
    dobvector <- filter(dobframe,yeardiagnosis <= years[ii])$dob
    if (length(dobvector) != 0) {
      # Make sure our dobvector isn't empty
      numcases[ii] <- removeDuplicates(dobvector,ignore)
    }
  }
  
  # If selected write final output to file
  if (!is.na(file)) {
    numcases[ii] <- removeDuplicates(dobvector,ignore, write = file)
  }
  
  # Return vector of unique cases
  return(numcases)
}

# Function to create file name for deduplicted output
deduplicateFile <- function(pop, state, mode, years) {
  filename <- paste(basePath, "/output/uniquehiv-", state, "_", mode, "_",
                    toString(tail(years,1)),sep="")
}

```

Removal of duplicates .....

```{r removeduplicates}

# Days to ignore in duplicate calculations
ignore <- c(1,15) 

# Store number unique diagnoses cumulatively and annually. For annual unique 
# cases assume all the first years of disgnoses are unique. 

### Do calculations for all notifications first
doball <- select(hivset,dob,yeardiagnosis) # Overall
numberUniqueAll <- numUnique(doball, allyears, ignore)

# Store national results
hivresults$uniquecases <- numberUniqueAll

# Add variable for proportion unique - due to statistical calculations 
# proportion mybe slightly higher than one. In those cases round down to 1.
hivresults$propunique <- hivresults$uniquecases/hivresults$totalnotifications
hivresults[hivresults$propunique > 1,]$propunique <- 1

### Now do states and territories
states <- c("act","nsw","vic","qld","nt","wa","sa","tas")

hivstate$uniquecases <- rep(NA,nrow(hivstate))
hivstate$propunique <- rep(NA,nrow(hivstate))

for (ii in 1:length(states)) {
  # Dob vector
  dobtemp <- select(subhivset(hivset, region = states[ii]), dob, yeardiagnosis)
  
  # Calculate number unique
  numberUnique <- numUnique(dobtemp, unique(dobtemp$yeardiagnosis), ignore)
  
  # Store state results
  hivstate[hivstate$state == states[ii],]$uniquecases <- numberUnique
}

# Calculate proportion unique
hivstate$propunique <- hivstate$uniquecases/hivstate$totalnotifications
hivstate[hivstate$propunique > 1,]$propunique <- 1

### Now do mode of exposure
modes <- c("msm", "pwid", "hetero", "other")

hivexpnat$uniquecases <- rep(NA,nrow(hivexpnat))
hivexpnat$propunique <- rep(NA,nrow(hivexpnat))

for (ii in 1:length(modes)) {
  # Dob vector
  dobtemp <- select(subhivset(hivset, mode = modes[ii]), dob, yeardiagnosis)
  
  # Calculate number unique
  numberUnique <- numUnique(dobtemp, unique(dobtemp$yeardiagnosis), ignore)
  
  # Store exposure results
  hivexpnat[hivexpnat$expgroup == modes[ii],]$uniquecases <- numberUnique
}

# Calculate proportion unique
hivexpnat$propunique <- hivexpnat$uniquecases/hivexpnat$totalnotifications
hivexpnat[hivexpnat$propunique > 1,]$propunique <- 1

# TODO: Country of births

```

```{r plotuniques}
graphics.off()

# Plot national notifications data raw and deduplicated
diagsplot1 <- ggplot(hivresults,aes(yeardiagnosis)) + theme_bw() +
  opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
  ylab("Number of notifications")
diagsplot1 <- diagsplot1 + geom_line(aes(y = totalnotifications,color="All notifications")) + 
  geom_line(aes(y = uniquecases,color="Unique notifications")) + 
  theme(legend.direction="vertical") + theme(legend.title=element_blank())

# diagsplot2 <- ggplot(hivexpnat,aes(yeardiagnosis,cumnotifications,fill = expgroup)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()
# 
# diagsplot3 <- ggplot(hivstate,aes(yeardiagnosis,cumnotifications,fill = state)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()
# 
# # Test by state
# test <- filter(hivexpstate,state == "nsw")
# 
# testplot <- ggplot(test,aes(yeardiagnosis,cumnotifications,fill = expgroup)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()

# Print figures
windows(width=6,height=5)
print(diagsplot1)
dev.print(png,paste(outputFolder, "/figures/unique_diagnoses.png",sep =""))

```

Sort out deathrates

```{r deaths}
# To calculate the number of deaths each year we calculate a deathrate with 
# an uncertainty range. This is used to work out the number of people
# diagnosed with HIV. 

# The estimate of the death rate uses a number of methods. Firstly uses the 
# 2003 linkage results and the number of known deaths to estimate the
# deathrate up to 2003. It then uses AHOD mortality rates from 2003 to  

# Load all the data sets we use for this estimate - use the parameter column as 
# row names for easier subsetting
hivparams <- read.csv(file.path(dataFolder, "individualHIVparameters.csv", 
                                sep =""),as.is = 1)
hivparams <- select(hivparams,parameter,value)
for (ii in 1:nrow(hivparams)) {
  assign(hivparams$parameter[ii],hivparams$value[ii])
}

# Load backprojection estimates for comparison
bpdeaths <- read.csv(file.path(dataFolder, "projectedDeaths.csv", 
                                sep =""),as.is = 1)

# Extract number of known deaths and store in a central data frame
hivdeaths <- hivdata %>% 
  filter(! is.na(yeardeath)) %>%
  group_by(yeardeath) %>% 
  summarise(number_deaths = n()) %>% 
  filter(yeardeath <= analysisYear)  

# Calculate sensitivity of recorded deaths based on linkage results
linkageFactor <- (linkedDeaths+notifiedDeaths)/notifiedDeaths
linkageFactorLower <- linkageFactor*(1-linkageSensitivity)
linkageFactorUpper <- linkageFactor*(1+linkageSensitivity)  

#  We don't need to adjust deaths for number of unique diagnoses as
# different dates of death should result in different people

# Set up linked deaths
hivdeaths$inflated_deaths <- hivdeaths$number_deaths * linkageFactor
hivdeaths$inflated_deaths_lower <- hivdeaths$number_deaths * linkageFactorLower
hivdeaths$inflated_deaths_upper <- hivdeaths$number_deaths * linkageFactorUpper

# No deaths in 1980 and 1982 so add this data manually to ensure everything lines up
hivdeaths <- rbind(hivdeaths,c(1980,0,0,0,0))
hivdeaths <- rbind(hivdeaths,c(1982,0,0,0,0))
hivdeaths <- arrange(hivdeaths,yeardeath)

### Now calculate mortality rate for linkage data ------------------------
cumdiagnoses <- hivresults$totalnotifications
deaths <- hivdeaths$inflated_deaths
deaths_min <- hivdeaths$inflated_deaths_lower
deaths_max <- hivdeaths$inflated_deaths_upper
anndiags <- c(cumdiagnoses[1],diff(cumdiagnoses))  # Annual diagnoses

# Calculate number living each year 
numliving <- rep(NA,length(anndiags))
numliving_min <- rep(NA,length(anndiags))
numliving_max <- rep(NA,length(anndiags))

numliving[1] <- anndiags[1]
numliving_min[1] <- anndiags[1]
numliving_max[1] <- anndiags[1]

for (ii in 2:length(anndiags)) {
  numliving[ii] <- numliving[ii-1] + anndiags[ii] - deaths[ii-1] 
  numliving_min[ii] <- numliving_min[ii-1] + anndiags[ii] - deaths_max[ii-1] 
  numliving_max[ii] <- numliving_max[ii-1] + anndiags[ii] - deaths_min[ii-1] 
}

# Calculate mortality
linkageMortality <- deaths/numliving
linkageMortalityMin <- deaths_min/numliving_max
linkageMortalityMax <- deaths_max/numliving_min

### Append AHOD mortality

# First find index corresponding to year of linkage 
indexLinkage <- max(seq(along = min(hivresults$yeardiagnosis):linkageYear))

# Save in deathrate
deathrate <- rep(NA,length(allyears))
deathrate_min <- rep(NA,length(allyears))
deathrate_max <- rep(NA,length(allyears))

deathrate[1:indexLinkage] <- linkageMortality[1:indexLinkage]
deathrate[(indexLinkage+1):length(deathrate)] <- ahodMortality

deathrate_min[1:indexLinkage] <- linkageMortalityMin[1:indexLinkage]
deathrate_min[(indexLinkage+1):length(deathrate)] <- ahodMortalityLower

deathrate_max[1:indexLinkage] <- linkageMortalityMax[1:indexLinkage]
deathrate_max[(indexLinkage+1):length(deathrate)] <- ahodMortalityUpper

# Store final death rate
hivdeathrate <- data.frame(year = allyears, deathrate = deathrate, 
                           deathrate_min = deathrate_min, 
                           deathrate_max = deathrate_max)

```

Sort out population movement rates

```{r migration}
# This chunk converts the ABS overseas and interste migration data into 
# population movement rates

# Load all the data sets we need - the ABS data is stored in a different folder
absFolder <- file.path(path.expand("~"),"Research","!Evaluation_Modelling","data")
 
absMigration <- read.csv(file.path(absFolder, "ABS_migration_clean.csv", 
                                sep =""),as.is = 1)
absInterstate <- read.csv(file.path(absFolder, "ABS_interstate_clean.csv", 
                                sep =""),as.is = 1)

# Calculate migration rate overall and for each state
absMigration <- mutate(absMigration, migrate = departures/erp)

states <- c("nsw", "vic", "qld", "nt", "wa", "sa", "tas", "act", "all")
nstates <- length(states)

allPredicts <- data.frame(year = allyears)

for (ii in 1:nstates) {
  mrate <- filter(absMigration, state == states[ii])
  
  # Now we need to fit and extrapolate for other years
  lmrate <- lm(migrate ~ year, data = mrate)
  expmrate <- predict(lmrate, data.frame(year = allyears))
  
  # Store extrapolated rate
  allPredicts[[states[ii]]] <- expmrate
}

# Now work out mins and maxs and store in a final data frame.
# Use state min and max as range for national migration rates.
# Use the relative range of the national rates for each state rate
allmrateMin <- apply(select(allPredicts,-year), 1, min)
allmrateMax <- apply(select(allPredicts,-year), 1, max)

# Calculate relative rates
relmrateMin <- allmrateMin/allPredicts$all
relmrateMax <- allmrateMax/allPredicts$all

# Store everything in a dataframe
hivmigrate <- gather(allPredicts, "state", "rate", 2:10)

hivmigrate$lower <- NA
hivmigrate$upper <- NA

hivmigrate[hivmigrate$state == "all",]$lower <- allmrateMin
hivmigrate[hivmigrate$state == "all",]$upper <- allmrateMax

for (ii in 1:(nstates-1)) {
  temprate <- filter(hivmigrate, state == states[ii])$rate
  hivmigrate[hivmigrate$state == states[ii],]$lower <-  relmrateMin * temprate
  hivmigrate[hivmigrate$state == states[ii],]$upper <-  relmrateMax * temprate
}

### Calculate interstate rates

# First estimate erp for each state

allinterstate <- data.frame(year = allyears)

for (ii in 1:nstates) {
  lmerp <- lm(erp ~ year, data = filter(absMigration, state == states[ii]))
  erpstate <- predict(lmerp, data.frame(year = allyears))
  
  # Store extrapolated population
  allinterstate[[states[ii]]] <- erpstate
}

hivinterstate <- gather(allinterstate, "state", "erp", 2:10)

# Calculate interstate ERP simply overall population minus state population
allerp <- filter(hivinterstate, state == "all")$erp

hivinterstate <- hivinterstate %>% 
  group_by(state) %>%
  mutate(intererp = allerp-erp)

# Calculate departure rate
hivinterstate$departures <- NA

for (ii in 1:nstates) {
  lmdepart <- lm(departures ~ year, data = filter(absInterstate, state == states[ii]))
  departstate <- predict(lmdepart, data.frame(year = allyears))
  
  # Store extrapolated population
  hivinterstate[hivinterstate$state == states[ii],]$departures <- departstate
}

hivinterstate <- mutate(hivinterstate,departrate = departures/erp) 

# Calculate arrival rate
hivinterstate$arrivals <- NA

for (ii in 1:nstates) {
  lmarrive <- lm(arrivals ~ year, data = filter(absInterstate, state == states[ii]))
  arrivestate <- predict(lmarrive, data.frame(year = allyears))
  
  # Store extrapolated population
  hivinterstate[hivinterstate$state == states[ii],]$arrivals <- arrivestate
}

hivinterstate <- mutate(hivinterstate,arriverate = arrivals/intererp) 

# As intererp is zero national set arriverate to departure rate for "all"
# Gives indication of rate people move interstate nationally
hivinterstate[hivinterstate$state == "all",]$arriverate <- 
  hivinterstate[hivinterstate$state == "all",]$departrate

```



```{r livingdiags}

# Now loop through the years to calculate the number living with diagnosed HIV
# using the linkage estimates and after the end of the linkage study the AHOD death rates

# Intialize results vectors
livingLinkage <- rep(0,length(allyears))
livingLinkageLower <- rep(0,length(allyears))
livingLinkageUpper <- rep(0,length(allyears))

livingAhod <- rep(0,length(allyears))
livingAhodLower <- rep(0,length(allyears))
livingAhodUpper <- rep(0,length(allyears))

deathsAhod <- rep(0,length(allyears))
deathsAhodLower <- rep(0,length(allyears))
deathsAhodUpper <- rep(0,length(allyears))

drateLinkage <- rep(0,length(allyears))
drateLinkageLower <- rep(0,length(allyears))
drateLinkageUpper <- rep(0,length(allyears))

# Set initial value
livingLinkage[1] <- hivresults$uniquecases[1]
livingLinkageLower[1] <- hivresults$uniquecases[1]
livingLinkageUpper[1] <- hivresults$uniquecases[1]

livingAhod[1] <- hivresults$uniquecases[1]
livingAhodLower[1] <- hivresults$uniquecases[1]
livingAhodUpper[1] <- hivresults$uniquecases[1]

# Loop through and produce all the calculations
for (ii in 1:(length(allyears)-1)){
  
  # Number living based on linkage 
  livingLinkage[ii + 1] <- livingLinkage[ii] + 
    hivresults$annualcases[ii] - 
    hivdeaths$inflated_deaths[ii] 
  livingLinkageLower[ii + 1] <- livingLinkageLower[ii] + 
    hivresults$annualcases[ii] - 
    hivdeaths$inflated_deaths_upper[ii] 
  livingLinkageUpper[ii + 1] <- livingLinkageUpper[ii] + 
    hivresults$annualcases[ii] - 
    hivdeaths$inflated_deaths_lower[ii] 
  
  # Calculate corresponding death rate
  drateLinkageLower[ii] <- hivdeaths$inflated_deaths_upper[ii]/
    livingLinkageLower[ii]
  drateLinkageUpper[ii] <- hivdeaths$inflated_deaths_lower[ii]/
    livingLinkageUpper[ii]
  
  # Calculate deaths for AHOD mortality rates
  if (allyears[ii] <= 2003) {
    deathsAhod[ii] <- hivdeaths$inflated_deaths[ii]
    deathsAhodLower[ii] <- hivdeaths$inflated_deaths_lower[ii]
    deathsAhodUpper[ii] <- hivdeaths$inflated_deaths_upper[ii]
  } else {
    deathsAhod[ii] <- ahodMortality * livingLinkage[ii]
    deathsAhodLower[ii] <- ahodMortalityLower * livingLinkageLower[ii]
    deathsAhodUpper[ii] <- ahodMortalityUpper * livingLinkageUpper[ii]
  }
  
  # Calculate living based on AHOD mortality
#   livingAhod[ii + 1] <- livingAhod[ii] + hivresults$annualcases[ii] - 
#     deathsAhod[ii] 
#   livingAhodLower[ii + 1] <- livingAhodLower[ii] + hivresults$annualcases[ii] - 
#     deathsAhodUpper[ii] 
#   livingAhodUpper[ii + 1] <- livingAhodUpper[ii] + hivresults$annualcases[ii] - 
#     deathsAhodLower[ii] 
}

# Insert final death values
finalIndex <- ii +1
drateLinkage[finalIndex] <- hivdeaths$inflated_deaths[finalIndex]/
  livingLinkage[finalIndex]
drateLinkageLower[finalIndex] <- hivdeaths$inflated_deaths_upper[finalIndex]/
  livingLinkageLower[finalIndex]
drateLinkageUpper[finalIndex] <- hivdeaths$inflated_deaths_lower[finalIndex]/
  livingLinkageUpper[finalIndex]
if (allyears[finalIndex] <= 2003) {
  deathsAhod[finalIndex] <- hivdeaths$inflated_deaths[finalIndex]
  deathsAhodLower[finalIndex] <- hivdeaths$inflated_deaths_lower[finalIndex]
  deathsAhodUpper[finalIndex] <- hivdeaths$inflated_deaths_upper[finalIndex]
} else {
  deathsAhod[finalIndex] <- ahodMortality * livingLinkage[finalIndex]
  deathsAhodLower[finalIndex] <- ahodMortalityLower * 
    livingLinkageLower[finalIndex]
  deathsAhodUpper[finalIndex] <- ahodMortalityUpper * 
    livingLinkageUpper[finalIndex]
}

# Store all the final death results
hivdeaths$deathsAhod <- deathsAhod
hivdeaths$deathsAhodLower <- deathsAhodLower
hivdeaths$deathsAhodUpper <- deathsAhodUpper

hivdeaths$drateLinkage <- drateLinkage
hivdeaths$drateLinkageLower <- drateLinkageLower
hivdeaths$drateLinkageUpper <- drateLinkageUpper

# Store all the number living with diagnosed HIV results 
# pldhiv <- data.frame(hivresults$yeardiagnosis,livingLinkage,livingLinkageLower,livingLinkageUpper,livingAhod,livingAhodLower,livingAhodUpper)
# names(pldhiv) <-c("year","livingLinkage","livingLinkageLower","livingLinkageUpper","livingAhod","livingAhodLower","livingAhodUpper")

# James's death file

# Plot deaths results for comparison
# deathsplot <- ggplot(hivdeaths,aes(x = yeardeath,y = inflated_deaths)) + geom_line()

```

Sort out treatment ....

Sort out treatment .....
