---
title: "Count HIV Cascade calculations"
author: "Richard T. Gray"
date: "Tuesday, April 21, 2015"
output: word_document
---

This Rmarkdown document describes an analysis of the HIV care and treatment cascade to estimate the proportion of new HIV infections in the overall population due to each stage of the cascade. The stages we are considering are people living with undiagnosed HIV (PLUHIV), people living with diagnosed HIV (PLDHIV), people receiving ART with detectable viral load, and people recieving ART with undetectable viral load. The analysis uses annual estimates for new infections and HIV notifications. 


```{r initialization,echo = FALSE,messages = FALSE,include=FALSE}
# Clear workspace
rm(list=ls()) 
options(scipen=999)  # To get rid of scientific notation

# Load libraries used
require(ggplot2)
require(reshape2)
require(RColorBrewer)
require(readxl)


# Details of where to store outputs
dataFolder <- "../data/"
outputFolder <- "../output/"
currTime <- format(Sys.time(), "%y-%m-%d") # to append to files
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
