---
title: "Count HIV Cascade calculations"
author: "Richard T. Gray"
date: "Tuesday, April 21, 2015"
output: word_document
---

This Rmarkdown document describes an analysis of the HIV care and treatment cascade to estimate the proportion of new HIV infections in the overall population due to each stage of the cascade. The stages we are considering are people living with undiagnosed HIV (PLUHIV), people living with diagnosed HIV (PLDHIV), people receiving ART with detectable viral load, and people recieving ART with undetectable viral load. The analysis uses annual estimates for new infections and HIV notifications. 

The results are written in an abstract format at the end of the document for submission to ASHM.....


```{r initialization, echo = FALSE, messages = FALSE, include=FALSE}
# Clear workspace
rm(list=ls()) 
options(scipen=999)  # To get rid of scientific notation

# Load libraries used and that nice for data manipulation
require(ggplot2)
# require(reshape2)
# require(RColorBrewer)
require(readxl)
require(dplyr)
require(tidyr)
require(utils)
require(nnls) # For optimization
require(knitr)
require(gridExtra)

# Set workin directory
setwd("C:/Users/Rgray/Documents/Research/!Evaluation_Modelling/projects/Care_Cascade_Calculations/code")

```

```{r inputdetails, echo = FALSE, messages = FALSE, include=FALSE}
# Details of where to store outputs
dataFolder <- "../data/"
outputFolder <- "../output/"
inputFile <- "countAbstractData.xlsx"
currTime <- format(Sys.time(), "%Y-%m-%d(%H-%M)") # to append to files

# Analysis parameters
numSamples <- 1000  # Number of sampled parameter sets for each year
numInits <- 1000    # Number of init starting points for optimization 

# Options for running and saving things
fullRun <- TRUE # Everytime we do a run from scratch save everything 
folder <- paste("CountAbstractAnalysis_",currTime,sep="")
if(fullRun){
  dir.create(file.path(outputFolder,folder), showWarnings = FALSE)
  outputFolder <- file.path(outputFolder,folder)
}

```


```{r loaddata, echo=FALSE, messages = FALSE, include=FALSE}
# Load data from the countdata.xlsx file 
countData <- read_excel(paste(dataFolder,inputFile,sep =""), na = "NA")

# Using the input data calculate numbers in each column
countClean <- countData[,c("year","diagnoses","infectsmin","infectsmax")]

countClean$undiagmin <- countData$diagnosedmin/(1-countData$undiagmin)-countData$diagnosedmin
countClean$undiagmax <- countData$diagnosedmax/(1-countData$undiagmax)-countData$diagnosedmax
countClean$diagnosedmin <- countData$diagnosedmin-countData$treatedmin
countClean$diagnosedmax <- countData$diagnosedmax-countData$treatedmax

# Number detectables is a little tricker to get the max and min right
countClean$detectablemin <- countData$treatedmin-countData$suppressedmax*countData$treatedmax
countClean$detectablemax <- countData$treatedmax-countData$suppressedmin*countData$treatedmin
countClean$undetectablemin <- countData$suppressedmin*countData$treatedmin
countClean$undetectablemax <- countData$suppressedmax*countData$treatedmax

# Save count files as csv files for later 
if (fullRun){
  write.csv(countData, file = file.path(outputFolder, paste(currTime, "_countRaw.csv", sep ="")))
  write.csv(countClean, file = file.path(outputFolder, paste(currTime, "_countFinal.csv", sep ="")))
}

# Display some output
print(head(countClean,n=2))

```

```{r sampling, echo=FALSE, messages = FALSE, include=FALSE}
# Produce a randomly sampled set of numbers for fitting from the ranges in the countClean dataframe. 

nyears <- nrow(countClean)
numRows <- nyears*numSamples # For preallocation

countSample <- data.frame(year = numeric(numRows),diagnoses = numeric(numRows), newinfects = numeric(numRows), 
                          undiagnosed = numeric(numRows), diagnosed = numeric(numRows),  detectable = numeric(numRows),
                          undetectable = numeric(numRows))

countAverage <- data.frame(year = numeric(nyears),diagnoses = numeric(nyears), newinfects = numeric(nyears),
                           undiagnosed = numeric(nyears), diagnosed = numeric(nyears), detectable = numeric(nyears),
                           undetectable = numeric(nyears))

for (ii in 1:nyears){
  startIndex <- (ii-1)*numSamples+1
  endIndex <- ii*numSamples
  # Fixed data
  countSample[startIndex:endIndex,]$year <- countClean$year[ii]*rep(1,numSamples)
  countSample[startIndex:endIndex,]$diagnoses <- countClean$diagnoses[ii]*rep(1,numSamples)
  # Random sampling
  countSample[startIndex:endIndex,]$newinfects <- runif(numSamples,countClean$infectsmin[ii], countClean$infectsmax[ii])
  countSample[startIndex:endIndex,]$undiagnosed <- runif(numSamples,countClean$undiagmin[ii],countClean$undiagmax[ii])
  countSample[startIndex:endIndex,]$diagnosed <- runif(numSamples,countClean$diagnosedmin[ii],countClean$diagnosedmax[ii])
  countSample[startIndex:endIndex,]$detectable <- runif(numSamples,countClean$detectablemin[ii],countClean$detectablemax[ii])
  countSample[startIndex:endIndex,]$undetectable <- runif(numSamples,countClean$undetectablemin[ii],countClean$undetectablemax[ii]) 
  
  countAverage[ii,] <- colMeans(countSample[startIndex:endIndex,])
}

# Save count sample and avearge as csv files for later 
if (fullRun){
  write.csv(countSample, file = file.path(outputFolder, paste(currTime, "_countSample.csv", sep ="")))
  write.csv(countAverage, file = file.path(outputFolder, paste(currTime, "_countAverage.csv", sep ="")))
}

# Display some output
print(head(countSample))
print(countAverage)

```

```{r exploration, echo=FALSE, messages = FALSE, include=FALSE}
# Before we get serious lets have a look at the data and what we are trying to do

# Melt data into a plotting data frame
datayear <- 2013
plotDataYear <- gather(filter(countSample, year == datayear),"indicator","estimate",4:7)
plotDataAve <- gather(countAverage,"indicator","estimate",4:7)

# Create a plot looking at the relationships
countplotYear <- ggplot(data=plotDataYear,aes(x=estimate,y=newinfects)) + geom_point() + 
  facet_wrap(~indicator,scales = "free_x") + theme_bw()

countplotAve <- ggplot(data=plotDataAve,aes(x=estimate,y=newinfects)) + geom_point() + 
  facet_wrap(~indicator,scales = "free_x") + theme_bw()

# Plot and save
windows(width=7,height=7)
print(countplotAve)
if (fullRun){
  ggsave(file.path(outputFolder, paste(currTime, "_CountPlot.png", sep ="")),width=7,height=7)
}

```

```{r analysis, echo=FALSE, messages = FALSE, include=FALSE}
# Now perform analysis trying to estimate beta values using a simple linear model

# Unadjusted analysis
#====================

# Try a standard linear regression first on everything without adjusting for known reductions when undetectable
# lm <- lm(diagnoses ~ 0 + undiagnosed + diagnosed + detectable + undetectable, data = countSample)
lm <- lm(newinfects ~ 0 + undiagnosed + diagnosed + detectable + undetectable, data = countSample)
print(lm)
coefUnadjustedOverall <- unname(coef(lm))

# Do the calculations for individual years
coefUnadjustedYear <- matrix(nrow = nrow(countClean),ncol = 4)
dataYears <- min(countSample$year):max(countSample$year)
for(ii in 1:length(dataYears)){
  lmYear <- lm(diagnoses ~ 0 + undiagnosed + diagnosed + detectable + undetectable, 
               data = filter(countSample,year == dataYears[ii]))
  coefUnadjustedYear[ii,] <- unname(coef(lmYear))
}
print(coefUnadjustedYear)

# Adjusted analysis
# ================

# Biogically we know undiagnosed people are more infectious due to acute infection or at least as infectious as diagnosed people. We also know undetectable PLHIV are 71 to 99% less infectious than diagnosed people. 

# Test forcing a constraint
countAdjust <- countSample
countAdjust$beta4 <- runif(numSamples,0.01,0.29)
countAdjust$adjusted <- (countSample$diagnosed + countAdjust$beta4*countSample$undetectable)

# Overall
# lmAdjust <- lm(diagnoses ~ 0 + undiagnosed + adjusted + detectable, data = countAdjust)
lmAdjust <- lm(newinfects ~ 0 + undiagnosed + adjusted + detectable, data = countAdjust)
print(lmAdjust)
coefAdjustedOverall <- unname(coef(lmAdjust))
confintTemp <- confint(lmAdjust)
coefLower <- c(confint(lmAdjust)[1,1],confint(lmAdjust)[2,1],confint(lmAdjust)[3,1])
coefUpper <- c(confint(lmAdjust)[1,2],confint(lmAdjust)[2,2],confint(lmAdjust)[3,2])

sampleUndiag <- runif(numSamples,coefLower[1],coefUpper[1])
sampleDiag <- runif(numSamples,coefLower[2],coefUpper[2])
sampleDetect <- runif(numSamples,coefLower[3],coefUpper[3])
sampleUndetect <- countAdjust$beta4*sampleUndiag

aveCoeffs <- colMeans(matrix(c(sampleUndiag,sampleDiag,sampleDetect,sampleUndetect),nrow = numSamples, ncol = 4))

numUndiag <- sampleUndiag*countSample$undiagnosed
numDiag <- sampleDiag*countSample$diagnosed
numDetect <- sampleDetect*countSample$detectable
numUndetect <- sampleUndetect*countSample$undetectable
numSum <- numUndiag + numDiag + numDetect + numUndetect

mean(numUndiag/numSum)
mean(numDiag/numSum)
mean(numDetect/numSum)
mean(numUndetect/numSum)

# Do the calculations for individual years
coefTemp <- matrix(nrow = nrow(countClean),ncol = 3)
coefAdjustedYear <- matrix(nrow = nrow(countClean),ncol = 4)

for(ii in 1:length(dataYears)){
  lmYearAdjusted <- lm(diagnoses ~ 0  + undiagnosed + adjusted + detectable, 
                       data = filter(countAdjust,year == dataYears[ii]))
  coefTemp[ii,] <- unname(coef(lmYearAdjusted))
}
coefAdjustedYear[,c(1,2,3)] <- coefTemp
coefAdjustedYear[,4] <- 0.05*coefTemp[,2]

print(coefAdjustedYear)

```

```{r}

# Constrained analysis
# ====================

countConstrain <- countAdjust
countConstrain$z <- -countConstrain$undiagnosed-countConstrain$adjusted

lmConstrain <- lm(z ~ 0 + undiagnosed + diagnoses + detectable, data = countConstrain)
print(lmConstrain)
coefs <- unname(coef(lmConstrain))

# Adjust the parameters
p2 <- -1/coefs[2]
p3 <- p2*coefs[3]
p1 <- (coefs[1] + 1)*p2

print(c(p1,p2,p3))
# Use Optim so we can add further constraints. Specifically that b is less than 0 and p2 < p1

# dat <- countConstrain[,c("z","undiagnosed","diagnoses","detectable")]

numInits <- 1000 #numSamples


# for(jj in 1:numYears)
dat <- tail(dat, n = numSamples) # Final year

min.RSS <- function(data,par){
  with(data,sum(par[1]*undiagnosed + par[2]*diagnoses + par[3]*detectable - z)^2)
}

results <- list()
values <- rep(numSamples,0)

initpars <- matrix(c(runif(numInits,0,10),runif(numInits,-50,0),runif(numInits,0,1)),nrow=numInits,ncol = 3)
for (ii in 1:numInits) {
  results[[ii]] <- optim(par = initpars[ii,], min.RSS, data = dat,lower=c(0,-50,0),
                         upper=c(10,0,1),method="L-BFGS-B")
  values[ii] <- results[[ii]]$value
}

# Convert parameters
# ==================

pars <- results[[which.min(values)]]$par
p2 <- -1/pars[2]
p3 <- p2*pars[3]
p1 <- (pars[1] + 1)*p2

# Final parameters
coefFinal <- c(p1,p2,p3,0.05*p2)
print(coefFinal)
```

```{r}
# Incidence result data frames
# =============================
coefUsed <- coefAdjustedOverall
# coefUsed <- coefFinal
# coefUsed <- tail(coefAdjustedYear, n = 1)
# coefUsed <- head(coefAdjustedYear, n = 1)

pldhivFrame <- countAverage[,c("year","undiagnosed","diagnosed","undetectable","detectable")] # For plotting
infectsFrame <- pldhivFrame
infectsFrame$undiagnosed <- coefUsed[1]*countAverage$undiagnosed
infectsFrame$diagnosed <- coefUsed[2]*countAverage$diagnosed
infectsFrame$detectable <- coefUsed[3]*countAverage$detectable
infectsFrame$undetectable<- coefUsed[4]*countAverage$undetectable

# Final year data
# propInfects <- infectsFrame[infectsFrame$year == 2013,2:5]/countAverage$diagnoses
propInfects <- infectsFrame[infectsFrame$year == 2013,2:5]/sum(infectsFrame[infectsFrame$year == 2013,2:5])
print(propInfects)

# Prepare to print out some results 
# propFrame <-data.frame(x=c("Undiagnosed","Diagnosed","On ART (detectable)","On ART (undetectable)"),y = 100*propInfects[1,])
# colnames(propFrame) <- c("Stage","Percentage")


```

```{r prettyplots, echo=FALSE, messages = FALSE, include=FALSE}
graphics.off()
# Melt data into right format 
# plhivResults <- gather(pldhivFrame,"stage","number",2:5)
plhivResults <- gather(pldhivFrame,"stage","number",2:5) %>%
                  dplyr::group_by(year) %>%
                    mutate(cumsum=cumsum(number))
infectsResults <- gather(infectsFrame,"stage","number",2:5) %>%
                    dplyr::group_by(year) %>%
                      mutate(cumsum=cumsum(number))

# Now create plots 

plhivPlotBase <- ggplot(data = plhivResults, aes(x=year,y=number, fill = stage)) +
  theme_bw() + xlab("Year") + labs(fill='Cascade stage') 
plhivPlot1 <- plhivPlotBase + geom_bar(stat="identity") + ylab("Number PLDHIV") 
plhivPlot2 <- plhivPlotBase + geom_bar(stat="identity",position="fill") + ylab("Proportion PLDHIV") 

infectsPlotBase <- ggplot(data = infectsResults, aes(x=year,y=number, fill = stage))+
  theme_bw() + xlab("Year") +  labs(fill='Cascade stage')
infectsPlot1 <- infectsPlotBase + geom_bar(stat="identity") + ylab("Number Diagnoses")
infectsPlot2 <- infectsPlotBase + geom_bar(stat="identity",position="fill") + ylab("Proportion Diagnoses")

# Print plots
windows(width=12,height=5)
print(grid.arrange(plhivPlot1,infectsPlot1,plhivPlot2,infectsPlot2,ncol = 2))

```

```{r tidyup,echo = FALSE}
options(scipen=0)  # Set back to default
```
