---
title: "HIV Cascades Calculations"
author: "Richard T. Gray"
date: Date - `r format(Sys.Date(), format="%B %d %Y")`
output: 
  word_document:
    reference_docx: mystyles.docx
---

This document describes and contains the code for producing the HIV care cascade estimates for Australia. It is written in dynamic format using R markdown v2 within Rstudio (version 0.98.1091 with R version`r substr(R.Version()$version.string,1,15)`) to produce a Word (.docx) document (see <http://rmarkdown.rstudio.com>). All the R code to clean the data and produce the figures and tables is contained in this file but suppressed in the final output document. 

```{r initialization, echo = FALSE, messages = FALSE, include=FALSE}
# Clear workspace
rm(list=ls()) 

# Libraries for data manipulation
require(readxl)
require(dplyr)
require(tidyr)
require(utils)

# Libraries for output
require(ggplot2)    
require(knitr)      
require(gridExtra)  
require(RColorBrewer)

# Set working directory
setwd(file.path(path.expand("~"),"Research","!Evaluation_Modelling","project_care_cascades","code"))

# Base directory
basePath <- file.path(path.expand("~"),"Research","!Evaluation_Modelling","project_care_cascades")
cleanDataFolder <- file.path(basePath,"data")

# Primary script parameters
currTime <- format(Sys.time(), "%y-%m-%d") # to append to files
analysisYear <- 2013
startYear <- 1990
# reclean <- FALSE  # Rerun data cleaning if true otherwise load already cleaned file.
pop <- "all"  # Population exposure category
region <- "all" # State for analysis

# TODO: Think about doing cumulative diagnoses for each state and population everytime

```

## Estimate cumulative HIV notifications 

```{r startAnalysis}
# Load clean file
hivdata <- read.csv(file.path(cleanDataFolder, paste("hivnotifications",toString(analysisYear),".csv", sep ="")))

# Setup overarching analysis                   
years <- startYear:analysisYear # Years to calculate and plot 
allyears <- min(hivdata$yeardiagnosis):analysisYear # All years of data

# Setup extra columns for analysis
hivdata$yeardiagnosis <- as.numeric(substr(hivdata$datediagnosis,1,4))
hivdata$yeardeath <- as.numeric(substr(hivdata$datedeath,1,4))

# Quick checks that we are not doing analysis outside of the data.
if (max(hivdata$yeardiagnosis) < analysisYear) {
  stop("No data for specified for final year of analysis.")
}

if (min(hivdata$yeardiagnosis) > startYear) {
  stop("No data for specified for first year of analysis.")
}

# Subset by population and state
if (pop == "all"){
  hivset <- hivdata
} else {
  hivset <- filter(hivdata, expgroup == pop)
}

if (region != "all") {
    hivset <- filter(hivset, state == region)
}
```

```{r plotcode, echo=FALSE,messages = FALSE,include=FALSE}
# Default plot specifications
graphics.off()

# Baseline theme for plot variables
opts <- theme(text = element_text(face = "bold",size=12,colour="black"),
  axis.text.x = element_text(face = "plain",size=10,colour="black"),
  axis.text.y = element_text(face = "plain",size=10,colour="black"),
  axis.line = element_line(colour="black"),
  axis.ticks = element_line(colour="black"),
  legend.position = "top",
  legend.background = element_rect(),
  legend.key = element_blank(),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"),
  plot.title=element_text(size=12, face="bold")
)

# Theme for pie chart
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  axis.text.y=element_blank(),
  plot.title=element_text(size=12, face="bold")
  )

# Setup colours
getPalette <- colorRampPalette(brewer.pal(12,"Set3"))
piecolours1 <- getPalette(12)
piecolours2 <- gray.colors(3)
piecolours <- c(piecolours2,piecolours1)
linecolours <- brewer.pal(4,"Set1")
```

Sort out diagnoses ...

```{r diagnoses}
# Setup dataframe for overall results by first extracting diagnoses details

# Overall annual results
hivresults <- hivset %>% 
  group_by(yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(totalnotifications  = cumsum(notifications)) #%>% 
  #filter(yeardiagnosis %in% years)   

# Calculate the proportion of diagnoses by exposure and population category by 
# state and nationally. Theses data frames are used for further calculations

hivstate <- hivset %>% group_by(state,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(cumnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(propcumnotifications = cumnotifications/sum(cumnotifications)) #%>%
  #filter(yeardiagnosis %in% years)

hivexpnat <- hivset %>% group_by(expgroup,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(cumnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(propcumnotifications = cumnotifications/sum(cumnotifications)) #%>%
  #filter(yeardiagnosis %in% years)

hivexpstate <- hivset %>% group_by(expgroup,state,yeardiagnosis) %>% 
  summarise(notifications = n()) %>% 
  mutate(cumnotifications = cumsum(notifications)) %>%
  ungroup() %>% 
  group_by(state,yeardiagnosis) %>% 
  mutate(propnotifications = notifications/sum(notifications)) %>%
  mutate(propcumnotifications = cumnotifications/sum(cumnotifications)) #%>%
  #filter(yeardiagnosis %in% years)

# TODO: Will have to do the same by state

```

Removal of duplicates .....

```{r removeduplicates}
# Load function to remove duplicates
source('~/Research/!Evaluation_Modelling/code/removeDuplicates.R', 
       echo=TRUE)

# Setup
ignore <- c(1,15) # Days to ignore in duplicate calculations
filename <- paste(basePath, "/output/uniquehiv-",pop,"_",region,"_",
                  toString(tail(years,1)),sep="")

# Set up dob frame
dobframe <- select(hivset,dob,yeardiagnosis)

# Now loop through years and calculate cumulative number of unique cases 
numUniques <- rep(0,length(allyears))
for (ii in seq(along=allyears)) {
  dobvector <- filter(dobframe,yeardiagnosis <= allyears[ii])$dob
  if (allyears[ii] == analysisYear) {
    numUniques[ii] <- removeDuplicates(dobvector,ignore,write=filename)
  } else {
    numUniques[ii] <- removeDuplicates(dobvector,ignore)
  }
}

# Store number unique diagnoses cumulatively and annually. For annual unique 
# cases assume all the first years of disgnoses are unique. 
hivresults$uniquecases <- numUniques
hivresults$annualcases <- c(numUniques[1],diff(numUniques))

# Plot national notifications data raw and deduplicated
diagsplot1 <- ggplot(hivresults,aes(yeardiagnosis)) + theme_bw() +
  opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
  ylab("Number of notifications")
diagsplot1 <- diagsplot1 + geom_line(aes(y = totalnotifications,color="All notifications")) + 
  geom_line(aes(y = uniquecases,color="Unique notifications")) + 
  theme(legend.direction="vertical") + theme(legend.title=element_blank())

# diagsplot2 <- ggplot(hivexpnat,aes(yeardiagnosis,cumnotifications,fill = expgroup)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()
# 
# diagsplot3 <- ggplot(hivstate,aes(yeardiagnosis,cumnotifications,fill = state)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()
# 
# # Test by state
# test <- filter(hivexpstate,state == "nsw")
# 
# testplot <- ggplot(test,aes(yeardiagnosis,cumnotifications,fill = expgroup)) + theme_bw() + 
#   opts + xlim(c(startYear,analysisYear)) + ylim(c(0,NA)) + xlab(" Year") + 
#   ylab("Number of notifications") + geom_area()

# Print figures
windows(width=6,height=5)
print(diagsplot1)
# dev.print(pdf,paste(outputFolder, "/figures/current_pie.pdf",sep =""))

```

Sort out deaths ....

```{r deaths}
# Now calculate the number of deaths each year to work out the number of people
# diagnosed with HIV 

# Load all the data sets we use for this estimate
hivparams <- read.csv(file.path(cleanDataFolder, "individualHIVparameters.csv", sep =""))
hivparams <- select(hivparams,parameter,value)

# James's death file
# AHOD death rates



```

Sort out treatment ....

Sort out treatment .....
