---
title: "Merging and Analysis of Australian ART Data"
author: "Richard T. Gray"
date: '`r format(Sys.Date(), format="%d %B %Y")`'
output: 
  word_document:
    pandoc_args: --output="docs/ART_analysis.docx"
---

This script explores the number of people taking antiretroviral terapy 
(ART) in Australia over 2000-2017. It estimates the number taking ART 
during each year by merging and fitting to two data sources. 

1. Australian HIV Observational Database (AHOD) which provides estimates 
for ART use up to 2012. The estimates over 2010-2012 are actually 
projections rather than estimates from raw data.
2. The Pharmaceutical Benefits Scheme (PBS) 10% longitudinal sample 
provided by the company Prospection. This data is only avaliable since
2013. It is considered a more reliable as it is a more direct estimate. 

The primary purpose of this analysis is to explore how to cover the 
change over period from the AHOD estimates to the PBS data. The results
are reported in the Australian HIV Cascade manuscript. 

```{r knitr_options, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  include = FALSE) 
```


```{r initialization} 
# Restart R and set working directory to source file location

# Various directories
basePath <- getwd() #dirname(getwd())
dataFolder <- file.path(basePath, "data")
#outputFolder <- file.path(basePath, "data")
figFolder <- file.path(basePath, "output","figures")
resultsFolder <- file.path(basePath, "output")
Rcode <- file.path(dirname(basePath), "code") 
HIVcode <- file.path(basePath, "code")

# Load standard libraries, key functions and options
source(file.path(Rcode, "LoadLibrary.R"), echo=TRUE)
source(file.path(Rcode, "DataLibraries.R"), echo=TRUE)
source(file.path(Rcode, "PlotOptions.R"), echo=TRUE)
source(file.path(Rcode, "TidyLongitudinal.R"), echo=TRUE)
source(file.path(Rcode, "AgeCat.R"), echo=TRUE)
LoadLibrary(cowplot)
LoadLibrary(captioner)


# Script parameters
analysisYear <- 2019
saveResults <- TRUE

# -------------------------------------------------------------------------

# Add temporary residents - Prior to 2018 we had two estimates one from the
# ATRAS study which has been published and one based on personal 
# communication with Aaron Cogle who did a survey for NAPWHA. In March 2019, 
# NAPHWA and KI produced a report providing more detailed estimates of the 
# number of temporary residents (Medicare ineligible) taking ART. The results 
# from this report were used in the 2018 estimates and updated for the 2019 
# estimates.

addTempResidents <- TRUE 
atrasTRest <- FALSE # TRUE if using the ATRAS estimate, False if using the 
                   # NAPWHA estimate. Use NAPWHA estimate for ASR and 
                   # ATRAS estimate for the paper as it is citable
if (addTempResidents) {
  tempResData <- tibble(population = NA, 
    year = 2013:analysisYear,
    value = NA,
    lower = NA,
    upper = NA)
  
  tempResData$population = "all"
  
  if (analysisYear <= 2017) { 
    if (atrasTRest) {
      # ATRAS estimate using a weighted calculation
      # Estimated 450 Medicare ineligible HIV+ temporary residents in 
      # Australia of which ~76% are on treatment (~340 people, calculated 
      # using a weighted proportion: 0.94*180 + 0.63*270 = 339.3. Assuming 
      # 63% of people not in ATRAS are on ART (percentage on ART at start 
      # of ATRAS) and 94% of people in ATRAS are on ART (percentage on ART 
      # at end of ATRAS). 
      # Range isassumed to be 300 to 400.
      # Citation: Petoumenos, Kathy, Jo Watson, Bill Whittaker, Jennifer Hoy, 
      # Don Smith, Lisa Bastian, Robert Finlayson, et al. "Subsidized Optimal 
      # ART for HIV-Positive Temporary Residents of Australia Improves
      # Virological Outcomes: Results from the Australian HIV Observational
      # Database Temporary Residents Access Study." Journal of the 
      # International AIDS Society 18, no. 1 (2015).
      
      # tempResData <- c(340, 300, 400)
      tempResData$value <- 340
      tempResData$lower <- 300
      tempResData$upper <- 400
      
    } else {
      # NAPWHA estimate from Aaron Cogle (used in 2017 ASR)
      # Citation: Personal communication, Aaron Cogle, National Association 
      # of People with HIV Australia, 7th July 2017. Survey of Medicare
      # ineligible people living with HIV and ART access, 2016. National
      # Association of People with HIV Australia (NAPWHA), Sydney. 
      # tempResData <- c(450, 500, 550)
      tempResData$value <- 450
      tempResData$lower <- 500
      tempResData$upper <- 550
    }
  } else if (analysisYear == 2018) {
    # Load file
    if (atrasTRest) {
      tempResData <- read.csv(file.path(dataFolder,
        "ART_medicare_ineligible-ATRAS.csv"))
    } else {
      tempResData <- read.csv(file.path(dataFolder,
        "ART_medicare_ineligible-NAPHWA.csv"))
    }
  } else {
  # Load from file ATRAS estimates merged with NAPHWA estimates in March 
  # 2019 report.
  tempResData <- read.csv(file.path(dataFolder,
    "ART_medicare_ineligible-2019.csv"))
  
  }
} else {
  tempResData$value <- 0
  tempResData$lower <- 0
  tempResData$upper <- 0
}


```

```{r loaddata}
# Load the ART data
artEstimates <- read.csv(file.path(dataFolder, "AHOD_ART_estimates.csv"))
prospectionData <- read.csv(file.path(dataFolder,
  paste("pharmdash_HIVpatients_clean-", toString(analysisYear),
    ".csv", sep = "")), as.is = c(1, 2))

# Clean ART data ---------------------------------------------------------

# First simplify column names
colnames(artEstimates) <- c("year", "ahod")

# Add national pbs data
tempPbs <- prospectionData %>% 
  filter(gender == "all", state == "all") %>%
  gather("year", "pbs", 3:ncol(prospectionData)) %>%
  select(year, pbs)

# Replace year strings with year values
tempPbs$year <- unlist(lapply(tempPbs$year, function(x)
  as.numeric(strsplit(as.vector(x), 'X')[[1]][2])))

# Add the PBS data to artEstimates
artEstimates <- full_join(artEstimates, tempPbs, by = "year")

# Add upper and lower bounds for the PBS data
if (analysisYear < 2019) {
popData <- read_excel(file.path(dirname(dirname(dataFolder)), "data",
                                "ABS_population_sizes.xlsx"),
                      sheet = 2)
} else {
  popData <- read_csv(file.path(dirname(dirname(dataFolder)), "data",
    "ABS_population_sizes_PharmDash-2019.csv"))
}

source(file.path(HIVcode, "pharmDashError.R"), echo=TRUE)

# Loop through years appending lower and upper PBS estimates
prospectionYears <- 2013:analysisYear 

for (nyear in prospectionYears) {
 
   numArt <- filter(artEstimates, year == nyear)$pbs
   numPop <- filter(popData, year == nyear, population == "all")$erp
   
   error <- PharmDashError(numPop, numArt)
   
   artEstimates$pbslower[artEstimates$year == nyear] <- error$lower
   artEstimates$pbsupper[artEstimates$year == nyear] <- error$upper
}

# Add temporary residents to PBS data
artEstimates$pbs[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbs[artEstimates$year %in% prospectionYears] +
  tempResData$value
artEstimates$pbslower[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbslower[artEstimates$year %in% prospectionYears] +
  tempResData$lower
artEstimates$pbsupper[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbsupper[artEstimates$year %in% prospectionYears] +
  tempResData$upper

# Add other variables ----------------------------------------------------

# Add some variables for specific data
artEstimates$ahodraw <- NA
artEstimates$all <- NA
artEstimates$ahodraw[1:11] <- artEstimates$ahod[1:11] # trust AHOD up to 2010
artEstimates$all[1:13] <- artEstimates$ahod[1:13]
artEstimates$all[14:nrow(artEstimates)] <- 
  artEstimates$pbs[14:nrow(artEstimates)]
```

## Fitting 

```{r Fit curves}
# Explore fitting a function to the data. Don't need to do the various
# logistic regressions because these can be done using ggplot with 
# geom_smooth. Rather we explore different fitting functions

# Functions to fit
fitFunc<- function(x, a, b, c, d) {
  # A four parameter logistic function
  return((b-a) / (1 + exp(-d * (x - c))) + a)
}

# Specify our x and y data for fitting
ydata <- artEstimates$all
xdata <- artEstimates$year

# Set weights to reflect PBS data is likely to be more accurate
# With first 11 AHOD estimates accurate, the next two (2011-2012) AHOD 
# estimates inaccurate.  
weights <- c(rep(0.66, 11), rep(0.33, 2), rep(1, length(prospectionYears)))

# Weighted logistic regression -------------------------------------------

# For comparison add a weighted regression fit
fitWlm <- lm(ydata ~ xdata, weights = weights)
artEstimates$fitwlm <- predict(fitWlm, xdata = xdata)

# GLM --------------------------------------------------------------------

# Try various general lineraized models

# First try a poisson regression
fitGlmPois <- glm(all ~ year, data = artEstimates, 
              family = poisson(link = "log"), weights = weights)
artEstimates$fitglmpois <- predict(fitGlmPois, type="response")

# Now try a binomial regression

saturationCoverage <- 30000 # estimate of saturation coverage

# Take the difference because we are taking the log
artEstimates$diffbin <- saturationCoverage - artEstimates$all 

# Now do a binomial fit
fitGlmBin <- glm(cbind(all, diffbin) ~ year, data = artEstimates, 
              family = binomial(link = "logit"), weights = weights)
artEstimates$fitglmbin <- predict(fitGlmBin, type="response") * 
  saturationCoverage

# Remove difference calculation
artEstimates <- select(artEstimates, -diffbin)

# Fit a specific function ------------------------------------------------

# Functions to fit (logistic)
fitFunc<- function(x, a, b, c, d) {
  # A four parameter logistic function
  return((b-a) / (1 + exp(-d * (x - c))) + a)
}

# Do the fitting 
if (analysisYear == 2014) {
  # Note for 2014 data: nls seems to have lots of problems fitting to
  # the full four parameter logistic function but works if we force one of 
  # of the parameters to a specific value. Here we set the inflection point.
  # For 2014 data it appears to be insensitive to a value greater than 2015. 
  inflectionYear <- 2020
  fitLog <- nls(ydata ~ fitFunc(xdata, a, b, inflectionYear, d),
  start = list(a = 5000, b = 30000, d = 0.1),
  weights = weights)
  
  # Note: best fit from Matlab for comparison for AnalysisYear = 2014. 
  # a = 5270, b = 30200, c = 2014.7 d = 0.207) 
  
} else if (analysisYear >= 2017) {
  # For 2017 data and after a four parameter fit can be done but confident 
  # intervals are problematic and mess up code below. We do a 4 parameter 
  # fit first and then fix the inflectionYear value and then refit
  fitLog4 <- nls(ydata ~ fitFunc(xdata, a, b, c, d),
    start = list(a = 5000, b = 30000, c = 2015, d = 0.1),
    weights = weights)
  
  inflectionYear <- coef(fitLog4)["c"]
  fitLog <- nls(ydata ~ fitFunc(xdata, a, b, inflectionYear, d),
  start = list(a = 5000, b = 30000, d = 0.1),
  weights = weights)

} else {
  fitLog <- nls(ydata ~ fitFunc(xdata, a, b, c, d),
    start = list(a = 5000, b = 30000, c = 2015, d = 0.1),
    weights = weights)
  
  inflectionYear <- coef(fitLog)["c"]
  warning("fitLog needs to beasssessed for Fit")
}

# Compare to an exponential
fitFuncExp <- function(x, a, b, c) {
  # A three parameter exponential function
  return(a * exp(b * (x - c)))
}

fitExp <- nls(ydata ~ fitFuncExp(xdata, a, b, 2000),
           start = list(a = 6233, b = 0.07),
           weights = weights)


# Add the fits to dataframe 
artEstimates$fitlog <- predict(fitLog, xdata = xdata)
artEstimates$fitexp <- predict(fitExp, xdata = xdata)

# Explore the fit - e.g logfit
# summary(fitLog)
# sum(resid(fitLog)^2)
# confint(fitLog)
```

```{r predictcurves}
# Generate ensemble of curves based on variation in 95% confidence interval
# of estimated parameters values
aRange <- confint(fitLog)[1,]
bRange <- confint(fitLog)[2,]
dRange <- confint(fitLog)[3,]

nsamples <- 10000  # ensure big enough to get enough samples out after 
                   # filtering
simMatrix <- matrix(0,nsamples, length(xdata))

aSample <- runif(nsamples, aRange[1], aRange[2])
bSample <- runif(nsamples, bRange[1], bRange[2])
dSample <- runif(nsamples, dRange[1], dRange[2])

for (ii in 1:nsamples) {
  simMatrix[ii,] <- fitFunc(xdata,aSample[ii], bSample[ii], 
                            inflectionYear, dSample[ii])
  
}

# Filter samples which go outside the PBS error bounds for 2017 
goodRows <- apply(simMatrix, 1,
      function(x) any(x[length(xdata)] <
                        artEstimates$pbsupper[length(xdata)]
                      && x[length(xdata)] >
                        artEstimates$pbslower[length(xdata)]))

simMatrix <- simMatrix[goodRows, ]


# Use function to generate lower and upper bounds for ART numbers for best 
# fitting function
artEstimates$fitmin <- apply(simMatrix, 2, min)
artEstimates$fitlwr95 <- apply(simMatrix, 2, quantile, probs = 0.025)
artEstimates$fitlwriqr <- apply(simMatrix, 2, quantile, probs = 0.25)
artEstimates$fitupriqr <- apply(simMatrix, 2, quantile, probs = 0.75)
artEstimates$fitupr95 <- apply(simMatrix, 2, quantile, probs = 0.975)
artEstimates$fitmax <- apply(simMatrix, 2, max)

# Add relative trends and adjusted values
finalIndex <- nrow(artEstimates)
index2012 <- nrow(filter(artEstimates, year <= 2012))

# Replace 2011:2012 values with fitted values as these were projected 
# anyway and are not direct estimates
artEstimates$alladjust <- artEstimates$all
artEstimates[artEstimates$year %in% c(2011:2012), ]$alladjust <- 
  artEstimates[artEstimates$year %in% c(2011:2012), ]$fitlog 

# Relative ART numbers
artEstimates$rel <- artEstimates$alladjust / 
  artEstimates$alladjust[finalIndex]

artEstimates$rellower <- artEstimates$fitmin / 
  artEstimates$alladjust

artEstimates$relupper <- artEstimates$fitmax / 
  artEstimates$alladjust
```


```{r plotanalysis}
figs <- captioner()

# Adjust plotOpts
plotOpts <- PlotOptions() + theme(legend.position = "right",
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9, face = "bold"))

#Setup colours, labels and shapes for legend aesthetics
cols <- c("ahod" = "blue", "pbs" = "red", "all" = "black", 
          "reg1" = "#99CCFF", "reg2" = "#663399", "reg3" = "#006699",
          "fitreg" = "#003300", "fitexp" = "#006699",
          "fitlog" = "#993333")

labels <- c("ahod" = "AHOD data", 
            "pbs" = "PBS data", 
            "all" = "ART data", 
            "reg1" = "Regression AHOD\n2000-2010",
            "reg2" = "Regression AHOD\n2000-2013", 
            "reg3" = "Regression all data", 
            "fitreg" = "Regression\nweighted", 
            "fitexp" = "Exponential fit",
            "fitlog" = "Logistic fit")

shapes <- c("ahod" = 18, "pbs" = 18, "all" = 18, 
            "reg1" = NA, "reg2" = NA, "reg3" = NA, 
            "fitreg" = NA, "fitexp" = NA, "fitlog" = NA)
            
# Create a plot for comparing the regression analysis
aesLabels <- c("ahod", "pbs", "fitreg", "reg3", "reg2", "reg1")

regressionPlot <- ggplot(data = artEstimates, aes(x = year, y = ahod)) + 
  geom_point(aes(colour = "ahod", shape = "ahod"), 
             na.rm = TRUE) + 
  geom_point(aes(y = pbs, colour = "pbs", shape = "pbs"), na.rm = TRUE) +
  geom_smooth(aes(y = ahodraw, colour = "reg1", shape = "reg1"),
              method = "lm", se = FALSE, na.rm = TRUE, fullrange = TRUE) + 
  geom_smooth(aes(colour = "reg2", shape = "reg2"), 
              method = "lm", se = FALSE, na.rm = TRUE, fullrange = TRUE) + 
  geom_smooth(aes(y = all, colour = "reg3", shape = "reg3"), 
              method = "lm",  se = FALSE, na.rm = TRUE, fullrange = TRUE) +
  geom_line(aes(y = fitwlm, colour = "fitreg", shape = "fitreg")) + 
  scale_colour_manual(name = "", values = cols,
                      labels = labels[aesLabels],
                      breaks = aesLabels,
                      guide = guide_legend(override.aes = list(
                         linetype = c("blank", "blank", 
                                      rep("solid", 4))))) +
  scale_shape_manual(name = "", values = shapes,
                     breaks = aesLabels, 
                     labels = labels[aesLabels]) +
  ylim(c(0,25000)) + 
  ylab("Number on ART") + 
  xlab("Year") + 
  plotOpts

fitPlot <- ggplot(data = artEstimates, aes(x = year)) +
  geom_point(aes(y = all, colour = "all", shape = "all"), na.rm = TRUE) +
  geom_line(aes(y = fitexp, colour = "fitexp", shape = "fitexp")) +
  geom_line(aes(y = fitlog, colour = "fitlog", shape = "fitlog")) +
  scale_colour_manual(name = "", values = cols,
                      labels = labels[c("all", "fitexp", "fitlog")],
                      breaks = c("all", "fitexp", "fitlog"),
                      guide = guide_legend(override.aes = list(
                         linetype = c("blank", "solid", "solid")))) +
  scale_shape_manual(name = "", values = shapes,
                     breaks = c("all", "fitexp", "fitlog"), 
                     labels = labels[c("all", "fitexp", "fitlog")]) +
  ylim(c(0,NA)) +
  ylab("Number on ART") +
  xlab("Year") +
  plotOpts

logPlot <- ggplot(data = artEstimates, aes(x = year)) + 
  geom_ribbon(aes(ymin = fitmin, ymax = fitmax, 
                  fill = "fitlog1", alpha = "fitlog1")) + 
  geom_ribbon(aes(ymin = fitlwr95, ymax = fitupr95,
                  fill = "fitlog2", alpha = "fitlog2")) +
  geom_ribbon(aes(ymin = fitlwriqr, ymax = fitupriqr,
                  fill = "fitlog3", alpha = "fitlog3")) +
  geom_point(aes(y = all, colour = "all", shape = "all"), na.rm = TRUE) +
  geom_errorbar(aes(ymin = pbslower, ymax = pbsupper), width = 0.3, 
                color = "black", size = 0.8, na.rm = TRUE) + 
  geom_line(aes(y = fitlog, colour = "fitlog", shape = "fitlog")) +
  scale_colour_manual(name = "Fit", values = cols, 
                      breaks = c("all", "fitlog"),
                      labels = labels[c("all", "fitlog")],
                      guide = guide_legend(override.aes = list(
                         linetype = c("blank", "solid")))) +
  scale_shape_manual(name = "Fit", values = shapes[c("all", "fitlog")], 
                     breaks = c("all", "fitlog"),
                     labels = labels[c("all", "fitlog")]) +
  scale_fill_manual(name = "Range",
                    values = c("#993333", "#993333", "#993333"),
                    breaks = c("fitlog1", "fitlog2", "fitlog3"),
                    labels = c("Range", "95% interval", "IQR")) +
  scale_alpha_manual(name =  "Range",
                    values = c(0.2, 0.35, 0.5),
                    breaks = c("fitlog1", "fitlog2", "fitlog3"),
                    labels = c("Range", "95% interval", "IQR")) +
  ylim(c(0,NA)) +
  ylab("Number on ART") +
  xlab("Year") +
  plotOpts

# Save plots as PNG files - 300 dpi by default
plotWidth <- 12
plotHeight <- 7
plotUnits <- "cm"

# Create a combined figure
fitsFigure <- ggdraw() +
    draw_plot(regressionPlot, 0, 0.5, 0.5, 0.5) +
    draw_plot(fitPlot, 0.5, 0.5, 0.5, 0.5) +
    draw_plot(logPlot, 0.2, 0, 0.6, 0.5) +
    draw_plot_label(c("A", "B", "C"), c(0, 0.5, 0.2), c(1, 1, 0.5),
      size = 12)

figs("fitsFigure", "(A) Linear regression fits to ART data. (B) Comparison 
of exponential and logistic curve fits to ART data. (C) Logistic function
fit with uncertainty.")


```

`r figs("fitsFigure")`
```{r Insert figure, include = TRUE, fig.width = 8, fig.height = 6}
fitsFigure
```

```{r Save results}
if (saveResults) {
 
  # String tag
  if (addTempResidents) {
    trString <- ifelse(atrasTRest,"ATRAS", "NAPWHA")
  } else {
    trString <- "PBS"
  }
  
  # Save plots
  ggsave(file.path(figFolder, paste0("Regression_Comparision-", 
    toString(analysisYear), "-", trString, ".png")), plot = regressionPlot,
    width = plotWidth, height = plotHeight, units = plotUnits)
  ggsave(file.path(figFolder, paste0("Fit_Comparison-", 
    toString(analysisYear), "-", trString, ".png")), plot = fitPlot, 
    width = plotWidth, height = plotHeight, units = plotUnits)
  ggsave(file.path(figFolder, paste0("Logistic_Fit-", 
    toString(analysisYear), "-", trString, ".png")), plot = logPlot,
    width = plotWidth, height = plotHeight, units = plotUnits)
  
  ggsave(file.path(figFolder, paste0("ART_Fitting_Figure-", 
    toString(analysisYear), "-", trString, ".png")), plot = fitsFigure,
    width = 8, height = 6, units = "in")
  
  # File directory and file name
  saveString <- file.path(resultsFolder, paste0("ART_Estimates-", 
    toString(analysisYear), "-", trString))
  
  # Write to csv
   write.csv(artEstimates, file = paste0(saveString, ".csv"), 
            row.names = FALSE)
}

```
