---
title: "Merging and Analysis of Australian ART data"
author: "Richard T. Gray"
date: '`r format(Sys.Date(), format="%d %B %Y")`'
output: 
  word_document:
    pandoc_args: --output="docs/ART_analysis.docx"
---

This script explores the number of people taking antiretroviral terapy 
(ART) in Australia over 2000-2014. It estimates the number taking ART during 
each year by merging and fitting to two data sources. 

1. Australian HIV Observational Database (AHOD) which provides estimates 
for ART use up to 2012. The estimates over 2010-2012 are actually 
projections rather than estimates from raw data.
2. The Pharmaceutical Benefits Scheme (PBS) 10% longitudinal sample 
provided by the company Prospection. This data is only avaliable since
2013. It is considered a more reliable as it is a more direct estimate. 

The primary purpose of this analysis is to explore how to cover the 
change over period from the AHOD estimates to the PBS data. The results
are reported in the Australian HIV Cascade manuscript. 

```{r knitr_options, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  include = FALSE) 
```

```{r Initialization} 
# Restart R and set working directory to source file location

# Various directories
basePath <- getwd() #dirname(getwd())
dataFolder <- file.path(basePath, "data")
#outputFolder <- file.path(basePath, "data")
figFolder <- file.path(basePath, "output","figures")
resultsFolder <- file.path(basePath, "output")
Rcode <- file.path(dirname(basePath), "code") 
HIVcode <- file.path(basePath, "code")

# Load libraries and key functions
source(file.path(Rcode, "LoadLibrary.R"))
LoadLibrary(tidyverse)
LoadLibrary(readxl)
LoadLibrary(scales)
LoadLibrary(cowplot)
LoadLibrary(captioner)
source(file.path(Rcode, "PlotOptions.R"), echo=TRUE)
source(file.path(Rcode, "TidyLongitudinal.R"), echo=TRUE)
source(file.path(Rcode, "AgeCat.R"), echo=TRUE)

# Script parameters
analysisYear <- 2014
saveResults <- FALSE

# Captions setup
figs <- captioner()

```

For this analysis we add available data on the number of temporary 
residents taking ART. Temporary residents are Medicare ineligible and hence
not counted in the PBS data. 

```{r Temporary residents}
# Add temporary residents - we have two estimates one from the ATRAS study 
# which has been published and one based on personal communication with 
# Aaron Cogle who did a survey for NAPWHA. 
addTempResidents <- FALSE
atrasTRest <- TRUE # TRUE if using the ATRAS estimate, False if using the 
# NAPWHA estimate. Use NAPWHA estimate for ASR and 
# ATRAS estimate for the paper as it is citable.
if (addTempResidents) {
  if (atrasTRest) {
    # ATRAS estimate using a weighted calculation
    # Estimated 450 Medicare ineligible HIV+ temporary residents in 
    # Australia of which ~76% are on treatment (~340 people, calculated 
    # using a weighted proportion: 0.94*180 + 0.63*270 = 339.3. Assuming 63% of
    # people not in ATRAS are on ART (percentage on ART at start of ATRAS) and 
    # 94% of people in ATRAS are on ART (percentage on ART at end of ATRAS). 
    # Range isassumed to be 300 to 400.
    # Citation: Petoumenos, Kathy, Jo Watson, Bill Whittaker, Jennifer Hoy, 
    # Don Smith, Lisa Bastian, Robert Finlayson, et al. "Subsidized Optimal 
    # ART for HIV-Positive Temporary Residents of Australia Improves
    # Virological Outcomes: Results from the Australian HIV Observational
    # Database Temporary Residents Access Study." Journal of the 
    # International AIDS Society 18, no. 1 (2015).
    
    tempResData <- c(340, 300, 400)
    
  } else {
    # NAPWHA estimate from Aaron Cogle (used in 2017 ASR)
    # Citation: Personal communication, Aaron Cogle, National Association 
    # of People with HIV Australia, 7th July 2017. Survey of Medicare
    # ineligible people living with HIV and ART access, 2016. National
    # Association of People with HIV Australia (NAPWHA), Sydney. 
    tempResData <- c(450, 500, 550)
  }
} else {
  tempResData <- c(0, 0, 0)
}


```

```{r Load data}
# Load the ART data
artEstimates <- read.csv(file.path(dataFolder, "AHOD_ART_estimates.csv"))
prospectionData <- read.csv(file.path(dataFolder,
  paste("pharmdash_HIVpatients_clean-", toString(analysisYear),
    ".csv", sep = "")), as.is = c(1, 2))

# Clean ART data ---------------------------------------------------------

# First simplify column names
colnames(artEstimates) <- c("year", "ahod")

# Add national pbs data
tempPbs <- prospectionData %>% 
  filter(gender == "all", state == "all") %>%
  gather("year", "pbs", 3:ncol(prospectionData)) %>%
  select(year, pbs)

# Replace year strings with year values
tempPbs$year <- unlist(lapply(tempPbs$year, function(x)
  as.numeric(strsplit(as.vector(x), 'X')[[1]][2])))

# Add the PBS data to artEstimates
artEstimates <- left_join(artEstimates, tempPbs, by = "year")

# Add upper and lower bounds for the PBS data
popData <- read_excel(file.path(dirname(dirname(dataFolder)), "data",
  "ABS_population_sizes_PharmDash.xlsx"), sheet = 2)
source(file.path(HIVcode, "pharmDashError.R"), echo=TRUE)

# Loop through years appending lower and upper pbs estimates
prospectionYears <- c(2013, 2014)

for (nyear in prospectionYears) {
  
  numArt <- filter(artEstimates, year == nyear)$pbs
  numPop <- filter(popData, year == nyear, population == "all")$erp
  
  error <- PharmDashError(numPop, numArt)
  
  artEstimates$pbslower[artEstimates$year == nyear] <- error$lower
  artEstimates$pbsupper[artEstimates$year == nyear] <- error$upper
}

# Add temporary residents to PBS data
artEstimates$pbs[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbs[artEstimates$year %in% prospectionYears] +
  tempResData[1]
artEstimates$pbslower[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbslower[artEstimates$year %in% prospectionYears] +
  tempResData[2]
artEstimates$pbsupper[artEstimates$year %in% prospectionYears] <-
  artEstimates$pbsupper[artEstimates$year %in% prospectionYears] +
  tempResData[3]

# Add other variables ----------------------------------------------------

# Add some variables for specific data
artEstimates$ahodraw <- NA
artEstimates$all <- NA
artEstimates$ahodraw[1:10] <- artEstimates$ahod[1:10]
artEstimates$all[1:13] <- artEstimates$ahod[1:13]
artEstimates$all[14:15] <- artEstimates$pbs[14:15]
```


## Fitting 

```{r Fit curves}
# Explore fitting a function to the data. Don't need to do the various
# logistic regressions because these can be done using ggplot with 
# geom_smooth. Rather we explore different fitting functions

# Specify our x and y data for fitting
ydata <- artEstimates$all
xdata <- artEstimates$year

# Set weights to reflect PBS data is likely to be more accurate
weights <- c(rep(0.66, 10), rep(0.33, 3), 1, 1)

# Weighted logistic regression -------------------------------------------

# For comparison add a weighted regression fit
fitWlm <- lm(ydata ~ xdata, weights = weights)
artEstimates$fitwlm <- predict(fitWlm, xdata = xdata)

# GLM --------------------------------------------------------------------

# Try various general lineraized models

# First try a poisson regression
fitGlmPois <- glm(all ~ year, data = artEstimates, 
  family = poisson(link = "log"), weights = weights)
artEstimates$fitglmpois <- predict(fitGlmPois, type="response")

# No try a binomial regression

saturationCoverage <- 30000 # estimate of saturation coverage

# Take the difference because we are taking the log
artEstimates$diffbin <- saturationCoverage - artEstimates$all 

# Now do a binomial fit
fitGlmBin <- glm(cbind(all, diffbin) ~ year, data = artEstimates, 
  family = binomial(link = "logit"), weights = weights)
artEstimates$fitglmbin <- predict(fitGlmBin, type="response") * 
  saturationCoverage

# Remove difference calculation
artEstimates <- select(artEstimates, -diffbin)

# Fit a specific function ------------------------------------------------

# Functions to fit (logistic)
fitFunc<- function(x, a, b, c, d) {
  # A four parameter logistic function
  return((b-a) / (1 + exp(-d * (x - c))) + a)
}

# Do the fitting - Note: nls seems to have lots of problems fitting to
# the full four parameter logistic function but works if we force one of 
# of the parameters to a specific value. Here we set the inflection point,
# it appears to be insensitive to a value greater than 2015. 
inflectionYear <- 2020

fitLog <- nls(ydata ~ fitFunc(xdata, a, b, inflectionYear, d),
  start = list(a = 5000, b = 30000, d = 0.1),
  weights = weights)

# Note: best fit from Matlab for comparison. 
# a = 5270, b = 30200, c = 2014.7 d = 0.207) 

# Compare to an exponential
fitFuncExp <- function(x, a, b, c) {
  # A three parameter exponential function
  return(a * exp(b * (x - c)))
}

fitExp <- nls(ydata ~ fitFuncExp(xdata, a, b, 2000),
  start = list(a = 6233, b = 0.07),
  weights = weights)


# Add the fits to dataframe 
artEstimates$fitlog <- predict(fitLog, xdata = xdata)
artEstimates$fitexp <- predict(fitExp, xdata = xdata)

# Explore the fit - e.g logfit
# summary(fitLog)
# sum(resid(fitLog)^2)
# confint(fitLog)
```

```{r Prediction curves}
# Generate ensemble of curves based on variation in 95% confidence interval
# of estimated parameters values
aRange <- confint(fitLog)[1,]
bRange <- confint(fitLog)[2,]
dRange <- confint(fitLog)[3,]

nsamples <- 10000  # ensure big enough to get enough samples out after 
# filtering
simMatrix <- matrix(0,nsamples, length(xdata))

aSample <- runif(nsamples, aRange[1], aRange[2])
bSample <- runif(nsamples, bRange[1], bRange[2])
dSample <- runif(nsamples, dRange[1], dRange[2])

for (ii in 1:nsamples) {
  simMatrix[ii,] <- fitFunc(xdata,aSample[ii], bSample[ii], 
    inflectionYear, dSample[ii])
  
}

# Filter samples which go outside the PBS error bounds for 2013 
goodRows <- apply(simMatrix, 1, 
  function(x) any(x[length(xdata)] <
      artEstimates$pbsupper[length(xdata)]
    && x[length(xdata)] >
      artEstimates$pbslower[length(xdata)]))

simMatrix <- simMatrix[goodRows, ]


# Use function to generate lower and upper bounds for ART numbers for best 
# fitting function
artEstimates$fitmin <- apply(simMatrix, 2, min)
artEstimates$fitlwr95 <- apply(simMatrix, 2, quantile, probs = 0.025)
artEstimates$fitlwriqr <- apply(simMatrix, 2, quantile, probs = 0.25)
artEstimates$fitupriqr <- apply(simMatrix, 2, quantile, probs = 0.75)
artEstimates$fitupr95 <- apply(simMatrix, 2, quantile, probs = 0.975)
artEstimates$fitmax <- apply(simMatrix, 2, max)

# Add relative trends and adjusted values
finalIndex <- nrow(artEstimates)
index2012 <- nrow(filter(artEstimates, year <= 2012))

# Replace 2012 value with fitted value
artEstimates$alladjust <- artEstimates$all
artEstimates[index2012, ]$alladjust <- artEstimates[index2012, ]$fitlog 

# Relative ART numbers
artEstimates$rel <- artEstimates$alladjust / 
  artEstimates$alladjust[finalIndex]

artEstimates$rellower <- artEstimates$fitmin / 
  artEstimates$alladjust

artEstimates$relupper <- artEstimates$fitmax / 
  artEstimates$alladjust
```

We explored numerous fitting options to the data over 2000-2014 including 
standard regresion approaches with weighting towards the PBS data. The most
plausible and best fit to the data was a four parameter logistic function.
This is what we use for our analysis in the Australian HIV cascade 
manuscript. 

## Results

```{r Plot analysis}
# Adjust plotOpts
plotOpts <- PlotOptions() + theme(legend.position = "right",
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9, face = "bold"))

#Setup colours, labels and shapes for legend aesthetics
cols <- c("ahod" = "blue", "pbs" = "red", "all" = "black", 
  "reg1" = "#99CCFF", "reg2" = "#663399", "reg3" = "#006699",
  "fitreg" = "#003300", "fitexp" = "#006699",
  "fitlog" = "#993333")

labels <- c("ahod" = "AHOD data", 
  "pbs" = "PBS data", 
  "all" = "ART data", 
  "reg1" = "Regression AHOD\n2000-2010",
  "reg2" = "Regression AHOD\n2000-2013", 
  "reg3" = "Regression all data", 
  "fitreg" = "Regression\nweighted", 
  "fitexp" = "Exponential fit",
  "fitlog" = "Logistic fit")

shapes <- c("ahod" = 18, "pbs" = 18, "all" = 18, 
  "reg1" = NA, "reg2" = NA, "reg3" = NA, 
  "fitreg" = NA, "fitexp" = NA, "fitlog" = NA)

# Create a plot for comparing the regression analysis
aesLabels <- c("ahod", "pbs", "fitreg", "reg3", "reg2", "reg1")

regressionPlot <- ggplot(data = artEstimates, aes(x = year, y = ahod)) + 
  geom_point(aes(colour = "ahod", shape = "ahod"), 
    na.rm = TRUE) + 
  geom_point(aes(y = pbs, colour = "pbs", shape = "pbs"), na.rm = TRUE) +
  geom_smooth(aes(y = ahodraw, colour = "reg1", shape = "reg1"),
    method = "lm", se = FALSE, na.rm = TRUE, fullrange = TRUE) + 
  geom_smooth(aes(colour = "reg2", shape = "reg2"), 
    method = "lm", se = FALSE, na.rm = TRUE, fullrange = TRUE) + 
  geom_smooth(aes(y = all, colour = "reg3", shape = "reg3"), 
    method = "lm",  se = FALSE, na.rm = TRUE, fullrange = TRUE) +
  geom_line(aes(y = fitwlm, colour = "fitreg", shape = "fitreg")) + 
  scale_colour_manual(name = "", values = cols,
    labels = labels[aesLabels],
    breaks = aesLabels,
    guide = guide_legend(override.aes = list(
      linetype = c("blank", "blank", 
        rep("solid", 4))))) +
  scale_shape_manual(name = "", values = shapes,
    breaks = aesLabels, 
    labels = labels[aesLabels]) +
  ylim(c(0,20000)) + 
  ylab("Number on ART") + 
  xlab("Year") + 
  plotOpts

fitPlot <- ggplot(data = artEstimates, aes(x = year)) +
  geom_point(aes(y = all, colour = "all", shape = "all"), na.rm = TRUE) +
  geom_line(aes(y = fitexp, colour = "fitexp", shape = "fitexp")) +
  geom_line(aes(y = fitlog, colour = "fitlog", shape = "fitlog")) +
  scale_colour_manual(name = "", values = cols,
    labels = labels[c("all", "fitexp", "fitlog")],
    breaks = c("all", "fitexp", "fitlog"),
    guide = guide_legend(override.aes = list(
      linetype = c("blank", "solid", "solid")))) +
  scale_shape_manual(name = "", values = shapes,
    breaks = c("all", "fitexp", "fitlog"), 
    labels = labels[c("all", "fitexp", "fitlog")]) +
  ylim(c(0,NA)) +
  ylab("Number on ART") +
  xlab("Year") +
  plotOpts

logPlot <- ggplot(data = artEstimates, aes(x = year)) + 
  geom_ribbon(aes(ymin = fitmin, ymax = fitmax, 
    fill = "fitlog1", alpha = "fitlog1")) + 
  geom_ribbon(aes(ymin = fitlwr95, ymax = fitupr95,
    fill = "fitlog2", alpha = "fitlog2")) +
  geom_ribbon(aes(ymin = fitlwriqr, ymax = fitupriqr,
    fill = "fitlog3", alpha = "fitlog3")) +
  geom_point(aes(y = all, colour = "all", shape = "all"), na.rm = TRUE) +
  geom_errorbar(aes(ymin = pbslower, ymax = pbsupper), width = 0.3, 
    color = "black", size = 0.8, na.rm = TRUE) + 
  geom_line(aes(y = fitlog, colour = "fitlog", shape = "fitlog")) +
  scale_colour_manual(name = "Fit", values = cols, 
    breaks = c("all", "fitlog"),
    labels = labels[c("all", "fitlog")],
    guide = guide_legend(override.aes = list(
      linetype = c("blank", "solid")))) +
  scale_shape_manual(name = "Fit", values = shapes[c("all", "fitlog")], 
    breaks = c("all", "fitlog"),
    labels = labels[c("all", "fitlog")]) +
  scale_fill_manual(name = "Range",
    values = c("#993333", "#993333", "#993333"),
    breaks = c("fitlog1", "fitlog2", "fitlog3"),
    labels = c("Range", "95% interval", "IQR")) +
  scale_alpha_manual(name =  "Range",
    values = c(0.2, 0.35, 0.5),
    breaks = c("fitlog1", "fitlog2", "fitlog3"),
    labels = c("Range", "95% interval", "IQR")) +
  ylim(c(0,NA)) +
  ylab("Number on ART") +
  xlab("Year") +
  plotOpts

# Save plots as PNG files - 300 dpi by default
plotWidth <- 12
plotHeight <- 7
plotUnits <- "cm"

# Create a combined figure
fitsFigure <- plot_grid(regressionPlot, fitPlot, logPlot,
  labels = c("A", "B", "C"), ncol = 1, nrow = 3)

figs("fitsFigure", "(A) Linear regression fits to ART data. (B) Comparison 
of exponential and logistic curve fits to ART data. (C) Logistic function
fit with uncertainty.")


```

`r figs("fitsFigure")`

```{r Insert figure, include = TRUE, fig.width = 5, fig.height = 9}
fitsFigure
```

```{r Save results}
if (saveResults) {
  
  # String tag
  if (addTempResidents) {
    trString <- ifelse(atrasTRest,"ATRAS", "NAPWHA")
  } else {
    trString <- "PBS"
  }
  
  # Save plots
  ggsave(file.path(figFolder, paste0("Regression_Comparision-", 
    trString, ".png")), 
    plot = regressionPlot,
    width = plotWidth, height = plotHeight, units = plotUnits)
  ggsave(file.path(figFolder, paste0("Fit_Comparison-", trString, ".png")), 
    plot = fitPlot, 
    width = plotWidth, height = plotHeight, units = plotUnits)
  ggsave(file.path(figFolder, paste0("Logistic_Fit-", trString, ".png")), 
    plot = logPlot,
    width = plotWidth, height = plotHeight, units = plotUnits)
  
  # File directory and file name
  saveString <- file.path(resultsFolder, paste0("ART_Estimates-", 
    toString(analysisYear), "-", trString))
  
  # Write to csv
  write.csv(artEstimates, file = paste0(saveString, ".csv"), 
    row.names = FALSE)
}

```
