Calculation for number of people living with diagnosed HIV
==========================================================

Neil Bretana and Richard T. Gray

This script is used for calculating the number of people 
living with diagnosed HIV in Australia. This is one of the key steps for
producing the estimates of the Australian HIV diagnosis and care cascade. 

This script is part of a project containing version 3.0 of the HIV cascade
calculation scripts. The main components of the script are in the 
CalculatePldhiv.R function. 

```{r Initialization}
# Chunk to setup everything

# Open as a project (setting working directory to source and restarting R)

# Setup directories
basePath <- getwd()
Rcode <- file.path(dirname(basePath), "code") # All cascades code
mainDataFolder <- file.path(dirname(getwd()), "data") # All cascades data
HIVcode <- file.path(basePath, "code") 
dataFolder <- file.path(basePath, "data")
outputFolder <- file.path(basePath, "output")
figuresFolder <- file.path(basePath, "output", "figures")

# Notifications folder is a private secure folder 
notificationsFolder <- file.path("/", "SVR-NAS", "Public", "SERP", "Data", 
  "National HIV Registry", "2017", "National", "Hamish & Richard")
# tempImputeFolder <- file.path(dirname(dirname(basePath)), "data", "HIV", 
#     "cascadeImputedHIVnotifications-2017")

# Load standard libraries and options ------------------------------------
source(file.path(Rcode, "LoadLibrary.R"))
LoadLibrary(tidyverse)
```

The following chunck specifies all the parameters for producing the
estimates of PLHIV previously diagnosed. Result files are saved to the
specified project folder and have a cascade file handle. 

*Current projects*
- ASR_2016, ASR_2017, ASR_2018 (cascades for the Annual Surveillance 
Report(ASR))
- ASR_2018_Ageing (age estimates of PLDHIV for ASR)
- CALD_Cascades
- HIV_Cascade_Paper
- Ageing_Paper

```{r Cascade and script parameters}
# Chunk to enter all the script parameters we will use
analysisYear <- 2017
saveResults <- TRUE

# Set HIV cascade parameters----------------------------------

# Overarching project folder where all cascades will be stored
projectOutput <- "Pldhiv_Mapping_2019" # "." for main output folder 
                                 # "HIV_Cascade_Paper" for paper

# Project CSV file for specifying cascade parameters - NULL = manual 
# otherwise cascade parameters are entred via a file in projectOutput 
# which need to exist
paramsFile <-"ProjectParameters.csv"
                            
# Cascade options----------------------------------------------------------
# The following variables specify the charactersitsics and hence the 
# notifications that will be included in the cascade. For some variables a 
# combination of categories can used by specifying a vector. Other 
# variables can only be specified as a single value.
# 
# When calculating multiple cascades these are entered as a list (of 
# vectors) for variables that can be combined or as vectors for single
# variables. All lists and vectors need to be the same length (equal to
# numCascades).
# 
# Note: These often have same value across all cascades in 
# a project so use rep() by default. These can then be manually changed as 
# required.

if (is.null(paramsFile)) {
  # Manually enter cascade specifications
  
  ##*********************************************************************##
  ## WARNING -- If multiple input categories
  ## are required you need to used the imputed notifications version.  
  ##*********************************************************************##  
  
  numCascades <- 1
  
  cascadeName <- "Test-2017" # (single)
  targetGender <- rep("all", numCascades) # all, male, female (single)
  targetAge <- rep("all", numCascades) # all or split (single)
  # Note regarding ageing: The calculations don't specifically target 
  # specific ages.  Rather this simply chooses whether age is considered or 
  # not."all" means there is no division by age whereas "split" means teh 
  # calculations are doen for all agegroups specified as a0_4, a5_9, 
  # a10_14,...,a85+ (5 year age bins). To estimate for specific age run with
  # split and then filter out age estimates as required essentially turns on 
  # doAge below. 
  targetCob <- rep("all", numCascades) # all=including n/a, Thailand, etc.
  # (combination). "Australia", "non-australia", "non-aus-nz"
  targetExposure <- rep("all", numCascades) #msm, hetero, pwid, otherexp 
  # (combination)
  targetAtsi <- rep("all", numCascades) # all, indigenous or non_indigenous. 
  # Only used if country is Australia (single)
  
  targetState <- rep(c("all"), numCascades) # nsw, sa, nt, qld, vic, wa, act
  # tas (combination)
  targetLocalRegion <- rep("all", numCascades) #tbd (combination)
  
  targetGlobalRegion <- rep("all", numCascades) # South-East Asia, 
  # Sub-Saharan Africa, Oceania, South American, Other cob, etc 
  # (combination)
  
  # Imputation options-------------------------------------------------------
  # These variables specify if the imputed notifications sets should be used 
  # and how many sets to use.
  # 
  useImputed <- rep(TRUE, numCascades)
  nImputedSets <- rep(10, numCascades)  # Default is 10, the maximum
  
  # Script options--------------------------------------------------------
  # The following variables specify the various calculation options. 
  # When calculating multiple cascades these are entered as a vector. All
  # vectors need to be the same length (equal to numCascades).
  # 
  # Note: These are generally given the same value across all cascades in 
  # a project so use rep() by default. These can then be manually changed  
  # as required. 
  excludeAborig <- rep(FALSE, numCascades) # FALSE be default to include 
                                           # indigenous
  doRetained <- rep(TRUE, numCascades) # TRUE for appending retained in care
  
  # Settings for doing deduplication analysis
  doUnique <- rep(TRUE, numCascades)
  yearUnique <- rep(2017, numCascades) #Use NULL or year 2100 to turn off
  saveUnique <- rep(FALSE, numCascades)
  
  # Settings for generating ECDC model inputs
  ecdcData <- rep(TRUE, numCascades)
  ecdcVersion <- rep("1.3.0", numCascades) # Has to be the same
  excludeOS <- rep(FALSE, numCascades) # Default FALSE but need to run with
                                       # TRUE for ASR as
  
  # Leave projections chunk turned off so it is not called unless required. 
  # Simply run projections from within the chunk. 
  projectPldhiv <- rep(FALSE, numCascades) 
  
  # Manually set final year for projections and projection name
  projectYear <- rep(analysisYear + 10, numCascades)  
  projectName <- rep("future_sq", numCascades)
  
  # Specifiy projection option for future diagnoses - can project 
  # under the current "status-quo" level or follow trend over last 5 years
  # or project a relative reduction
  projectOption <- rep("status-quo", numCascades) #status-quo, linear, reduce 
  projectDecrease <- 1 # 1 (0% reduction), 0.65 (35% reduction), 0.5 (50% 
                          # reduction), 0.35 (65% reduction) 
                          # relative reduction!
  
} else {
  # Read in a CSV file in 'projectOutput' (note project output folder has 
  # to already exist) with the 'paramsFile' name. 
  
  cascadeParameters <- read_csv(file.path(outputFolder, projectOutput,
    paramsFile))
  
  # Number of cascades
  numCascades <- nrow(cascadeParameters)
  
  # Pull out all the parameters separately. Default values are shown 
  # after each parameter (use in file and change as required)
  cascadeName <- cascadeParameters$cascadeName # Default: "All-analysisYear"
  targetGender <- cascadeParameters$targetGender # Default: "all"
  targetAge <- cascadeParameters$targetAge # Default: "all"
  targetCob <- cascadeParameters$targetCob # Default: "all"
  targetExposure <- cascadeParameters$targetExposure # Default: "all"
  targetAtsi <- cascadeParameters$targetAtsi # Default: "all"
  targetState <- cascadeParameters$targetState # Default: "all"
  targetLocalRegion <- cascadeParameters$targetLocalRegion # Default: "all"
  targetGlobalRegion <- cascadeParameters$targetGlobalRegion # Default: "all"
  
  # Imputation options
  useImputed <- cascadeParameters$useImputed # Default: TRUE
  nImputedSets <- cascadeParameters$nImputedSets # Default: 10
  
  # Script options
  excludeAborig <- cascadeParameters$excludeAborig # Default: FALSE
  doRetained <- cascadeParameters$doRetained # Default: TRUE
  
  # Deduplication analysis
  doUnique <- cascadeParameters$doUnique # Default: TRUE
  yearUnique <- cascadeParameters$yearUnique # Default: 2017
  saveUnique <- cascadeParameters$saveUnique # Default: FALSE
  
  # ECDC model
  ecdcData <- cascadeParameters$ecdcData # Default: TRUE
  ecdcVersion <- cascadeParameters$ecdcVersion # Default: "1.3.0"
  excludeOS <- cascadeParameters$excludeOS # Default: FALSE
  
  # Projections - defaults set to status quo for 10 years 
  projectPldhiv <- cascadeParameters$projectPldhiv # Default: FALSE
  projectYear <- cascadeParameters$projectYear # Default: 2017
  projectName <- cascadeParameters$projectName # Default: analysisYear + 10
  projectOption <- cascadeParameters$projectOption # Default: "future_sq"
  projectDecrease <- cascadeParameters$projectDecrease # Default: 0
    
} 

```

```{r Functions}
# Load functions used for analysis 

# Define simple but useful functions for looking at things
viewDiagnosed <- function() {
  print(pldhivAll %>% tail())
  print(pldhivAllMin %>% tail())
  print(pldhivAllMax %>% tail())
}

# Functions for setting up vectors for projections
End <- function(vector) return(tail(vector, 1))
ProjVec <- function(vector, nyears) return(c(vector, 
  rep(End(vector), nyears)))

# Functions used by CalculatePldhiv.R -------------------------------------
# Source functions for filling in missing data
source(file.path(Rcode, "FillMissing.R"))
source(file.path(Rcode, "FillDataFrame.R"))
source(file.path(Rcode, "FillMissingAge.R"))

# Load function to calculate number living with diagnosed HIV
source(file.path(HIVcode,"LivingDiagnosed.R"))
source(file.path(HIVcode,"LivingDiagnosedAge.R"))

# Function to easily extract sub-populations of interest 
source(file.path(HIVcode, "SubHivSet.R"))
source(file.path(HIVcode, "SubHivSetImpute.R"))

# Functions for removing duplicates annually
source(file.path(HIVcode, "DeduplicationFunctions.R"))

# Function to replace estimates
source(file.path(HIVcode, "ReplaceEstimates.R"))

# Function to tidy notifications
# source(file.path(HIVcode, "TidyNotifications.R"))

# Functions to output ECDC model data
source(file.path(HIVcode, "EcdcCalculations.R"))
source(file.path(HIVcode, "EcdcFiles.R"))

# Functions for subsetting and annual notifications
source(file.path(HIVcode, "ProportionMale.R"))
source(file.path(HIVcode, "AnnualDiagnoses.R"))

# Functions for adjustments
source(file.path(HIVcode, "GetAdjustments.R"))
source(file.path(HIVcode, "GetMigrate.R"))
source(file.path(HIVcode, "GetInterRegion.R"))

# Function for getting state/territory from local region
source(file.path(HIVcode, "RegionState.R"))

# Perform and append retained in care calculations
if (any(doRetained)) {
  source(file.path(HIVcode, "RetainedCare.R"))
}

# For producing future estimates of diagnoses for projections
if (any(projectPldhiv)) {
  source(file.path(HIVcode, "ProjectDiagnoses.R"))
}
```

```{r Load notifications}
# This chunk loads the national notifications subset and imputed sets 
# based on the specified criteria in Script paramaters

# Load cascade notifications data from storage
hivData <- read_csv(file.path(notificationsFolder, 
  paste0("cascadeImputedHIVnotifications-", toString(analysisYear)),
  "cascadeHIVdata.csv"), col_types = cols(
    .default = col_character(),
    yeardiagnosis = col_double(),
    agehiv = col_integer(),
    dob = col_date(format = ""),
    yearaids = col_double(),
    yeardeath = col_double(),
    dateill = col_date(format = ""),
    previ_diag_overseas = col_integer(),
    postcode = col_integer()
  ))

# Load imputed notifications sets if required
if (any(useImputed)) {
  hivDataImputed <- read_csv(file.path(notificationsFolder, 
     paste0("cascadeImputedHIVnotifications-", toString(analysisYear)),
    "cascadeHIVimputed.csv"), col_types = cols(
    .default = col_character(),
    yeardiagnosis = col_double(),
    agehiv = col_integer(),
    dob = col_date(format = ""),
    yearaids = col_double(),
    yeardeath = col_double(),
    dateill = col_date(format = ""),
    previ_diag_overseas = col_integer()))
}

# Data checks-------------------------------------------------------------

# Quick check that we are not doing analysis outside of the data
if (max(hivData$yeardiagnosis) < analysisYear) {
  stop("No data specified for final year of analysis.")
}

# HIV data variables------------------------------------------------------

# Setup variables for overarching analysis                  
allYears <- min(hivData$yeardiagnosis):analysisYear # all years of data

```

```{r Load base and adjustment data}
# This loads the adjustment files for adjusting the death rates and
# migration rates  
hivBase <- read.csv(file.path(dataFolder, paste0("HIVbaseEstimates-",
  toString(analysisYear), ".csv")))
hivAdjustments <- read.csv(file.path(dataFolder,
  paste0("HIVadjustments-",
    toString(analysisYear), ".csv")))

# Load cleaned NOM data variable (cleanNom)
load(file = file.path(mainDataFolder, "cleaned_ABS_nom_data.R"))

# Interstate movement
absInterstate <- read_csv(file.path(mainDataFolder, 
  paste0("ABS_interstate_age_sex_clean-", toString(analysisYear), ".csv")),
  col_types = cols(
    year = col_integer(),
    region = col_character(),
    sex = col_character(),
    age = col_character(),
    type = col_character(),
    value = col_integer()
  ))

# Inter-region movement
absInterRegion <- read_csv(file.path(mainDataFolder, 
  paste0("ABS_interregion_age_sex_clean-", toString(analysisYear), ".csv")),
  col_types = cols(
    year = col_integer(),
    region = col_character(),
    sex = col_character(),
    age = col_character(),
    arrivals = col_integer(),
    departures = col_integer(),
    erp = col_integer()
  ))

# Specifications and deaths data by age only used if doAge = TRUE ---------

# Set-up age matrix names
ageList <- c("a0_4", "a5_9","a10_14", "a15_19", "a20_24", "a25_29", 
  "a30_34", "a35_39", "a40_44", "a45_49", "a50_54", "a55_59", 
  "a60_64", "a65_69", "a70_74", "a75_79", "a80_84", "a85+")

yearList <- paste0("y", as.character(allYears))

# Load age based relative deathrates 
hivAgeDeath <- read.csv(file.path(dataFolder, paste0("HIVageEstimates-",
  toString(analysisYear), ".csv")))

# Load region codes
regionCodes <- read_csv(file.path(dataFolder, "postcodeRegionCodes.csv"),
  col_types = cols(
    postcode = col_integer(),
    SA3_code = col_integer(),
    SA3_name_2016 = col_character()
))
```

```{r Loop through cascades}
# This chunk loops through the cascade parameters and produces the 
# PLDHIV cascade estimates by calling the main pldhiv calculation function.
# Everything is pretty much done in that function.
source(file.path(HIVcode, "CalculatePldhiv.R"))

# List of lists to store output
pldhivCalculations <- list()

# Vector to store original number of notifications if imputation is used
numNotificationsOrig <- c() 

for (ii in 1:numCascades) {
  
  # Select correct data set
  if (useImputed[ii]) {
    pldhivData <- hivDataImputed
    
    numNotificationsOrig[ii] <- nrow(SubHivSetImpute(hivData, "all", targetGender, 
      targetExposure, targetCob, targetAtsi, targetState, 
      targetLocalRegion, targetGlobalRegion)[[1]])
    
  } else {
    pldhivData <- hivData
  }
  
  # Make sure targetState matches targetLocalRegion - targetState should
  # be set to "all" (and will be changed here) or the correct state
  # corresponding to the local region. If not an error is produced. 
  if (targetLocalRegion[ii] != "all") {
    if (targetState[ii] == "all") {
      # Change state to match targetLocalRegion 
      targetState[ii] <- RegionState(targetLocalRegion[ii], regionCodes)
    } else {
      # Check state matches target region
      currentState <- RegionState(targetLocalRegion[ii], regionCodes)
      if (currentState != targetState[ii]) {
        stop("Error: targetState does not align with targetLocalRegion") 
      }
    }
  }
  
  print(cascadeName[ii])
  
  # Do the calulations - note uniaue notifications output is at the 
  # state level not local region level. 
  pldhivCalculations[[ii]] <- CalculatePldhiv(analysisYear, saveResults, 
    projectOutput, 
    # Cascade parameters
    cascadeName[ii], targetGender[ii], targetAge[ii], targetCob[ii], 
    targetExposure[ii], targetAtsi[ii], targetState[ii], 
    targetLocalRegion[ii], targetGlobalRegion[ii], 
    # Specifications
    useImputed[ii], nImputedSets[ii], excludeAborig[ii], 
    doRetained[ii], doUnique[ii], yearUnique[ii], ecdcData[ii], 
    ecdcVersion[ii], excludeOS[ii], projectPldhiv[ii], projectYear[ii], 
    projectName[ii], projectOption[ii], projectDecrease[ii], 
    # Data
    pldhivData, allYears, hivBase, hivAdjustments, cleanNom,
    absInterstate, absInterRegion, ageList, yearList, hivAgeDeath)
  
}


```

