---
title: "HIV Cascade by Age Results"
author: "Richard T. Gray"
date: Latest version - `r format(Sys.Date(), format="%B %d %Y")`
output:
  word_document:
    pandoc_args: --output="docs/HIVcascade_Age.docx"
    reference_docx: docs/mystyles.docx
csl: docs/plos.csl
---

This document is used to generate and explore the Australian HIV cascade
by age and trends in the age distribution of people living with diagnosed
HIV over time. It produces the age distribution results for the Annual 
Surveillance Report and all the results and figures for the manuscript
Bretana et al, "An ageing population of people living with diagnosed HIV in 
Australia".

```{r knitr_options, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  include = FALSE) 
```

```{r Setup}
# Open as a project (setting working directory to source and restarting R)

# Setup directories
basePath <- getwd() #dirname(getwd())
Rcode <- file.path(dirname(getwd()), "code")
HIVcode <- file.path(basePath, "code") 
dataFolder <- file.path(basePath, "data")
resultsFolder <- file.path(basePath, "output")

# Load libraries and standard functions
source(file.path(Rcode, "LoadLibrary.R"))
LoadLibrary(tidyverse)
LoadLibrary(cowplot)
LoadLibrary(captioner)
LoadLibrary(scales)
LoadLibrary(RColorBrewer)
LoadLibrary(gridExtra)
source(file.path(Rcode, "FormatData.R"))
source(file.path(Rcode, "PlotOptions.R"))
source(file.path(Rcode, "PlotColors.R"))
source(file.path(Rcode, "SaveFigure.R"))
source(file.path(Rcode, "GetLegend.R"))
source(file.path(HIVcode, "ResultsFunctions.R"))

# Script parameters
savePlots <- TRUE

# Set up project and cascade to analyse
projectName <- "Ageing_Paper" #"ASR_2018_Ageing"  
resultsFolder <- file.path(resultsFolder, projectName)
cascadeNames <- c("All", "Male", "Female")

asr <- FALSE # TRUE = numbers for ASR report; FALSE = paper

# Create output directory
if (savePlots) {
  saveFormat <- ".png"
  figFolder <- file.path(resultsFolder, "figures")
  dir.create(figFolder, showWarnings = FALSE, recursive = TRUE)
}

# Captions
figs <- captioner()
# tabs <- captioner(prefix = "Table")

```

```{r Load results}
# This chunk loads all the results we need and stores them in long form 
# data frames

baseFile <- "HIVpldhivEstimates-2017-age"

# Load main results up to 2017
allMain <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, ".csv"))) %>%
  mutate(set = "all")
maleMain <- read_csv(file.path(resultsFolder, "Male-2017",
  paste0(baseFile, ".csv"))) %>%
  mutate(set = "male")
femaleMain <- read_csv(file.path(resultsFolder, "Female-2017",
  paste0(baseFile, ".csv"))) %>%
  mutate(set = "female")

mainResults <- bind_rows(allMain, maleMain, femaleMain) %>%
  select(set, everything())

# Load future projection results for main HIV cascade
if (!asr) {
sqResults <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, "-future_sq.csv"))) %>%
  mutate(set = "sq")

linearResults <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, "-future_linear.csv"))) %>%
  mutate(set = "linear")
linearResultsMale <- read_csv(file.path(resultsFolder, "Male-2017",
  paste0(baseFile, "-future_linear.csv"))) %>%
  mutate(set = "linear")
linearResultsFemale <- read_csv(file.path(resultsFolder, "Female-2017",
  paste0(baseFile, "-future_linear.csv"))) %>%
  mutate(set = "linear")

reduce35Results <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, "-future_reduce35.csv"))) %>%
  mutate(set = "reduce35")
reduce50Results <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, "-future_reduce50.csv"))) %>%
  mutate(set = "reduce50")
reduce65Results <- read_csv(file.path(resultsFolder, "All-2017",
  paste0(baseFile, "-future_reduce65.csv"))) %>%
  mutate(set = "reduce65")

futureResults <- bind_rows(sqResults, linearResults, reduce35Results,
  reduce50Results, reduce65Results) %>%
  select(set, everything())
}

```

```{r Calculate average ages}
# This chunk is used to calculate the average age of PLDHIV for 
# specific years

# Main results
AverageAge(allMain, 2007)
AverageAge(allMain, 2017)
AverageAge(maleMain, 2007)
AverageAge(maleMain, 2017)
AverageAge(femaleMain, 2007)
AverageAge(femaleMain, 2017)

# Projection results
if (!asr) {
  pldhivLinear <- linearResults %>%
    rename(value = pldhiv) 
  pldhivLinearMale <- linearResultsMale %>%
    rename(value = pldhiv) 
  pldhivLinearFemale <- linearResultsFemale %>%
    rename(value = pldhiv) 
  
  AverageAge(pldhivLinear, 2017)
  AverageAge(pldhivLinear, 2027)
  
  pldhivReduce35 <- reduce35Results %>%
    rename(value = pldhiv) 
  AverageAge(pldhivReduce35, 2017)
  AverageAge(pldhivReduce35, 2027)
  
  pldhivReduce50 <- reduce50Results %>%
    rename(value = pldhiv) 
  AverageAge(pldhivReduce50, 2017)
  AverageAge(pldhivReduce50, 2027)
} 
```


```{r Plot age trends}
# This chunk produces age trend plots for each main cascade

# Only produce results for > 15 year olds
ages <- unique(allMain$agebin)
ageBins <- ages[4:length(ages)]

# Use PlotAgeCascades function with default start year of 1986 and default
# colours and labels
allPlots <- PlotAgeCascade(allMain, agebins = ageBins)
malePlots <- PlotAgeCascade(maleMain, agebins = ageBins)
femalePlots <- PlotAgeCascade(femaleMain, agebins = ageBins)

# Plot up to 2017 for males and females
if (!asr) {
  all2027plots <- PlotAgeCascade(pldhivLinear, agebins = ageBins, 
    startyear = 1987)
  male2027plots <-PlotAgeCascade(pldhivLinearMale, agebins = ageBins, 
    startyear = 1987)
  female2027plots <-PlotAgeCascade(pldhivLinearFemale, agebins = ageBins, 
    startyear = 1987)
}

# Save plots
if (savePlots) {
  
  # Standard dimensions
  width <- 17.5
  height <- 13
  units <- "cm"
  
  # Save all plots
  SaveFigure(figFolder, "All-2017_number", allPlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "All-2017_proportion", allPlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Male-2017_number", malePlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Male-2017_proportion", malePlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Female-2017_number", femalePlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Female-2017_proportion", femalePlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  
  if (!asr) {
    SaveFigure(figFolder, "Male-2027_number", male2027plots[[1]],
      format = saveFormat, width = width, height = height, units = units)
    SaveFigure(figFolder, "Female-2027_number", female2027plots[[1]],
      format = saveFormat, width = width, height = height, units = units)
  }
  
}
```

```{r ASR results and plots}
# For the ASR we produce PLDHIV estimates for < 50 and 50+ age groups.
if (asr) {
  asrAges <- list(c("a0_4", "a5_9","a10_14", "a15_19", "a20_24", "a25_29",
    "a30_34", "a35_39", "a40_44", "a45_49"), c("a50_54", "a55_59", "a60_64",
      "a65_69", "a70_74", "a75_79", "a80_84", "a85+"))
  asrNames <- c("< 50", "50 +")
  
  # ASR estimates
  allAsr <- MergeAgeCascade(allMain, asrAges, asrNames)
  maleAsr <- MergeAgeCascade(maleMain, asrAges, asrNames)
  femaleAsr <- MergeAgeCascade(femaleMain, asrAges, asrNames)
  
  # ASR plots
  allAsrPlots <- PlotAgeCascade(allAsr, startyear = 1986, 
    agebins = asrNames, plotcolors = PlotColors("hivcascade4")[1:2], 
    grayscale = FALSE, agenames = asrNames)
  maleAsrPlots <- PlotAgeCascade(maleAsr, startyear = 1986, 
    agebins = asrNames, plotcolors = PlotColors("hivcascade4")[1:2], 
    grayscale = FALSE, agenames = asrNames)
  femaleAsrPlots <- PlotAgeCascade(femaleAsr, startyear = 1986, 
    agebins = asrNames, plotcolors = PlotColors("hivcascade4")[1:2], 
    grayscale = FALSE, agenames = asrNames)
  
  # Save the ASR estimates in he figures folder
  write_csv(allAsr, file.path(figFolder, "pldhiv-2017-ASR-age-all.csv"))
  write_csv(maleAsr, file.path(figFolder, "pldhiv-2017-ASR-age-males.csv"))
  write_csv(femaleAsr, file.path(figFolder, 
    "pldhiv-2017-ASR-age-female.csv"))
  
  # Save the ASR plots
  SaveFigure(figFolder, "All-2017_ASR_number", allAsrPlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "All-2017_ASR_proportion", allAsrPlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Male-2017_ASR_number", maleAsrPlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Male-2017_ASR_proportion", maleAsrPlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Female-2017_ASR_number", femaleAsrPlots[[1]],
    format = saveFormat, width = width, height = height, units = units)
  SaveFigure(figFolder, "Female-2017_ASR_proportion", femaleAsrPlots[[2]],
    format = saveFormat, width = width, height = height, units = units)
  
}
```

```{r Paper plot}
# Create a combined plot for paper
if (!asr) {
  legend <- GetLegend(male2027plots[[1]])
  paperFigureA <- male2027plots[[1]] + labs(title = "Males") +
    theme(legend.position = "none")
  paperFigureB <- female2027plots[[1]] + labs(title = "Females") +
    theme(legend.position = "none")
  
  paperFigure <- ggdraw() +
    draw_plot(paperFigureA, 0, 0, 0.455, 1) +
    draw_plot(paperFigureB, 0.455, 0, 0.455, 1) +
    draw_plot(legend, 0.91, 0, 0.09, 1) +
    draw_plot_label(c("A", "B"), c(0, 0.5),
      c(1, 1), size = 12)
  
  figs("paperfig", "Estimated number of diagnosed males (A) and females (B)
    in Australia over 1987-2027 by 5-year age bin (under projection 
    scenario).")
  
  if (savePlots) {
    SaveFigure(figFolder, "Combined_paper_figure-2027", paperFigure,
      format = saveFormat, width = 27, height = 10, units = units)
  }
}
```


```{r Plot age distribution}
# This chunk produces a simple age distrubtion for the number of people
# diagnosed with HIV in 2016 overall and by sex and overall at the end
# of the projections in 2016. 
if (!asr) {
  resultsYear <- 2017
  currentSets <- c("all", "male", "female")
  currentSetNames <- c("Overall", "Males", "Females")
  ageDist2016 <- PlotAgeDist(mainResults, resultsYear, sets = currentSets,
    setnames = currentSetNames)
  
  futureYear <- 2027
  futureDists <- futureResults %>%
    rename(value = pldhiv)
  futureSets <- unique(futureResults$set)
  futureSetNames <- c("Status-quo", "Linear", "Reduce by 35%", 
    "Reduce by 50%", "Reduce by 65%")
  ageDist2026 <-PlotAgeDist(futureDists, futureYear, sets = futureSets,
    setnames = futureSetNames)
  
  if (savePlots) {
    
    # Standard dimensions
    width <- 12
    height <- 9
    units <- "cm"
    
    # Save plots
    SaveFigure(figFolder, "Distribution-2017", ageDist2016,
      format = saveFormat, width = width, height = height, units = units)
    SaveFigure(figFolder, "All-Distribution-2027", ageDist2026,
      format = saveFormat, width = width, height = height, units = units)
    
  }
} 
```

```{r Plot future projections}
# This chunk produces simple plots of the overall number of people
# living with diagnosed HIV. 
if (!asr) {
  startYear <- 2012
  endYear <- 2027
  futureSets <- unique(futureResults$set)
  futureSetNames <- c("Status-quo", "Linear", "Reduce by 35%", 
    "Reduce by 50%", "Reduce by 65%")
  futurePldhiv <- futureResults %>% 
    group_by(year, set) %>%
    summarise(pldhiv = sum(pldhiv))
  
  projectPlot <- PlotPldhivProjection(futurePldhiv, startYear, endYear, 
    sets = futureSets, setnames = futureSetNames)
  
  if (savePlots) {
    
    # Save plots
    SaveFigure(figFolder, "Pldhiv_projections-2017", projectPlot,
      format = saveFormat, width = 12, height = 8)
  }
}
```

