---
title: "HIV diagnosis and care cascade for Australia"
resource_files:
- HIVcascadeEstimates-2014.csv
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: https://github.com/leftygray/Cascade_calculations/dashboards
    theme: simplex
    vertical_layout: fill
---
    
<!-- Learn more about flexdashboard at https://rstudio.github.io/flexdashboard -->    
    
```{r setup, include = FALSE}
library(dplyr)
library(knitr)
library(tidyr)
library(readr)
library(ggplot2)
library(flexdashboard)
library(shiny)
library(maps)
library(plotly)
library(treemap)

# Years of validity - to initialize start and end years
startYear <- 2005 
endYear <- 2014

# Load HIV cascade data
inputFile <- file.path(paste("HIVcascadeEstimates-", 
  toString(endYear), ".csv", sep = ""))
hivCascade <- read_csv(inputFile)

# Load Australia map data by state - mising ACT!!
ozdata <- read_csv("ozdata.csv")

```

```{r tidy up data, include = FALSE}    
# Exclude linked to care if it exists
hivCascade <- filter(hivCascade, !(stage == "linked"))

# Plot specs
stages <- c("infected", "pldhiv",  
            "retained", "numART", "suppressed")
stageNames <- c("Living with HIV", "Diagnosed", 
                  "Retained in care", "Receiving ART", "Suppressed virus")
stageNamesNeat <- c("Living \n with HIV", "Diagnosed", 
                      "Retained \n in care", 
                      "Receiving \n ART", "Suppressed \n virus")

# Final results we want
resultsAll <- filter(hivCascade,
                     population == "all")

cascadeYear <- reactive({
  resultsAllYear <- filter(resultsAll, year == input$resultsYear)
  return(resultsAllYear)
})

```

```{r plot options, include = FALSE}
plotOpts <- theme_bw() + theme(text = element_text(face = "bold", 
                                  size = 12, colour = "black"),
                               axis.text.x = element_text(face = "plain",
                                  size = 10, colour = "black"),
                               axis.text.y = element_text(face = "plain",
                                  size = 10, colour = "black"),
                               axis.line = element_line(colour = "black"),
                               axis.ticks = element_line(colour = "black"),
                               legend.position = "top",
                               legend.background = element_rect(),
                               legend.key = element_blank(),
                               panel.grid.major = element_blank(), 
                               panel.grid.minor = element_blank(), 
                               panel.background = element_blank(), 
                               panel.border = element_rect(colour = "black"),
                               axis.line = element_line(colour = "black"),
                               plot.title=element_text(size = 12, 
                                                       face="bold"),
                               strip.background = element_blank())

# Colour scheme used in ASR
asrcols <- c("#621C20", "#AB2322", "#A63603", "#D95728", "#E97164")

```
        

Annual cascade
==========================================================================

Inputs {.sidebar}
--------------------------------------------------------------------------

This dashboard provides an interface for exploring the HIV diagnosis and 
care cascade for Australia. On this tab you can explore the cascade for 
each year from 2003 by jurisdiction and population. The change in the 
cascade estimates over time for each satge can be explored on the 
**Change over time** tab. A detailed description of the methods used to 
generate these estimates is available on the **About** tab.

```{r}    
numericInput("resultsYear", "Select year", value = 2014,
    min = startYear, max = endYear)

selectInput("jurisdiction", "Select jurisdiction (not working yet)",
            choices = c("All", "ACT", "NSW", "NT", "QLD", "SA", 
                        "TAS","VIC", "WA"), 
            selected = "All")


selectInput("population", "Select population (not working yet)",
            choices = c("All", "bornsea", "bornssa", "hetero",
                        "indigenous", "msm", 
                        "non-indigenous","othercob", "otherexp",
                        "pwid"), 
            selected = "All")

```

**Things to do**

- Put ACT on the map!
- Implement jursidiction and population selections (requires a 
restructuring of the input data file)
- Add button to download CSV data file
- Fix small initial rendering issue when you change tabs

Application author: [Richard T. Gray](http://kirby.unsw.edu.au/people/dr-richard-gray), 
[The Kirby Instiute, UNSW Australia](http://kirby.unsw.edu.au/)

row 
--------------------------------------------------------------------------

### People undiagnosed {.value-box}
```{r}
renderValueBox({
  nUndiagnosed <- filter(cascadeYear(), stage == "infected")$value -
    filter(cascadeYear(), stage == "pldhiv")$value
  valueBox(
    value = round(nUndiagnosed, digits = -1),
    color = asrcols[1]
  )
})
```

### Diagnosed people not on ART {.value-box}
```{r}
renderValueBox({
  nUntreated <- filter(cascadeYear(), stage == "pldhiv")$value -
    filter(cascadeYear(), stage == "numART")$value
  valueBox(
    value = round(nUntreated, digits = -1),
    color = asrcols[2]
  )
})
```

### People with unsuppressed virus {.value-box}
```{r}
renderValueBox({
  nUnsuppressed <- filter(cascadeYear(), stage == "numART")$value -
    filter(cascadeYear(), stage == "suppressed")$value
  valueBox(
    value = round(nUnsuppressed, digits = -1),
    color = asrcols[5]
  )
})
```

row 
--------------------------------------------------------------------------

### Number of PLHIV by Jurisdiction

```{r}
# Produce a map with each state and territory coloured based on number
# of PLHIV

states <- c("all", "nsw", "vic", "qld", "nt", "wa", "sa", "tas", "act")

renderPlotly({
  mapData <- hivCascade %>%
    filter(population %in% tail(states, -1), stage == "infected",
           year == input$resultsYear) %>%
    select(population, value) %>%
    rename(state = population)
  
  ozdata$state <- tolower(ozdata$state)
 
  
  ozplhiv <- right_join(ozdata, mapData, by = "state")
  ozplhiv$text <- with(ozplhiv, paste("State: ", toupper(state),
                                    " Number: ", round(value)))
  
  mapPlot <- ggplot(data = ozplhiv, 
                    aes(x = long, y = lat, group = state)) +
    geom_polygon(aes(fill = value, text = text)) + 
    scale_fill_gradient(name = "Number of\n PLHIV", 
                        low = asrcols[5], high = asrcols[1]) +
    theme_void()

  ggplotly(mapPlot, tooltip = "text")
})
```

### HIV Diagnosis and Care Cascade 
    
```{r}

# Plot a bar chart for the results year 
renderPlot({ggplot(data = cascadeYear(),
                   aes(x = stage, y = value)) +
  geom_bar(stat = "identity", color = asrcols,
           fill = asrcols, width = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1,
                color = "black", size = 1.1) +
  scale_x_discrete(limits = stages, labels = stageNamesNeat) +
  ylab("Number of people") + xlab("") +
  plotOpts
})
```

row 
--------------------------------------------------------------------------
    
### Number of PLHIV by exposure   

```{r}
riskpops <- c("all", "msm", "hetero", "pwid", "otherexp")

renderPlot({
  riskData <- hivCascade %>%
    filter(population %in% tail(riskpops, -1), stage == "infected",
           year == input$resultsYear) %>%
    select(population, value) %>%
    rename(group = population)

  
  riskPlot <- treemap(riskData, index = c("group"),
                      vSize = "value", vColor = "value",
                      palette = asrcols, title = "Mode of exposure")
  
  riskPlot
  # ggplotly(mapPlot, tooltip = "text")
})

```

### HIV Cascade Estimates and Range
    
```{r}
# Produce a table of the results year cascade
renderTable({
  tableResults <- cascadeYear() %>%
  select( -year, -population) %>%
  rename(Stage = stage, Estimate = value, 
         Lower = lower, Upper = upper) %>%
  slice(c(2, 1, 4, 3, 5))

  tableResults$Stage <- stageNames
  
  tableResults
  })
```

Change over time
===========================================================================


Inputs {.sidebar}
--------------------------------------------------------------------------

This tab is used to explore the change in the cascade and 
individual satges over time. 

```{r}    
renderUI({numericInput("startStageYear", "Select start year", 
                       value = input$resultsYear - 10 +1,
                       min = 2004, max = input$resultsYear)})

stageOptions <- c("Living with HIV" = "infected", 
                "Diagnosed" = "pldhiv", 
                "Receiving ART" = "numART", 
                "Suppressed virus" = "suppressed")

selectInput("stage", "Select cascade stage",
            choices = names(stageOptions), 
            selected = "All")

```

row 
---------------------------------------------------------------------------


### PLHIV Over Time

```{r}
plotLabels <- c("Undiagnosed", 
                  "Diagnosed untreated", 
                  "On ART:  VL > 400 last test", 
                  "On ART: VL < 400 last test")
  
plotCols <- asrcols[c(1, 2, 4, 5)]

# Set up data for plotting
stackResults <- resultsAll %>%
  filter(stage != "retained") %>%
  select(year, stage, value) %>%
  spread(stage, value) %>%
  mutate(undiagnosed = infected - pldhiv,
         diagosed = pldhiv - numART,
         unsuppressed = numART - suppressed) %>%
  select(year, undiagnosed, diagosed, unsuppressed, suppressed) %>%
  gather("stage", "value", 2:5)

renderPlot({
  # Create plots
  ggplot(data = filter(stackResults, year >= input$startStageYear),
                          aes(x = year, y = value,fill = stage)) +
    xlab("Year") +
    scale_fill_manual(values = plotCols, name = "", labels = plotLabels, 
                      guide = guide_legend(reverse=TRUE, nrow = 2)) +  
    scale_x_continuous(breaks = c(input$startStageYear,
                                  input$resultsYear)) + 
    plotOpts + 
    theme(legend.text = element_text(size = 8)) +
    geom_bar(stat="identity") + 
    ylab("Number of PLHIV") 
})

```

### Proportional Breakdowm  of PLHIV

```{r}

renderPlot({
   
  ggplot(data = filter(stackResults, year >= input$startStageYear), 
                          aes(x = year, y = value, fill = stage)) +
    xlab("Year") +
    scale_fill_manual(values = plotCols, name = "", labels = plotLabels, 
                      guide = guide_legend(reverse=TRUE, nrow = 2)) +  
    scale_x_continuous(breaks = c(input$startStageYear,
                                  input$resultsYear)) + plotOpts + 
    theme(legend.text = element_text(size = 8)) +
    geom_bar(stat="identity", position= "fill") + 
    ylab("Number of PLHIV") 
})
```

row
---------------------------------------------------------------------------

### HIV Cascade Stage Over Time

```{r}
                  


yLabels <- c("infected" = "Number of PLHIV",
             "pldhiv" = "Number diagnosed",
             "numART" = "Number on ART",
             "suppressed" = "Number with VL < 400")

plotStages <- c("infected", "pldhiv", "numART", "suppressed")

# Plot the stage over time
renderPlot({
  resultsYears <- resultsAll %>%
    filter(year >= input$startStageYear)
  
  plotData <- filter(resultsYears, stage %in% plotStages)
  
  stageData <- filter(plotData, stage == stageOptions[input$stage])
  
  ggplot(data = stageData, aes(x = year, y = value)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              fill = asrcols[2], alpha = 0.4) +
  geom_line(color = asrcols[2]) +
  scale_x_continuous(breaks = c(input$startStageYear,
                                  input$resultsYear),
                     limits = c(input$startStageYear,
                                  input$resultsYear)) +
  expand_limits(y = 0) +
  ylab(yLabels[stageOptions[input$stage]]) + xlab("Year") +
  plotOpts
})

```

### HIV Cascade Stage Estimates and Range

```{r}

renderTable({
  stageResults <- resultsAll %>%
    filter(stage == stageOptions[input$stage], 
           year >= input$startStageYear,
           year <= input$resultsYear) %>%
    select(-population, -stage) %>%
    rename(Estimate = value, Lower = lower, Upper = upper, Year = year)
  
  stageResults
})

```

About
=====

The following sections describe the methods and data sources used to produce the estimates for each stage of the cascade.

**Estimating the number of people with diagnosed infection**

To estimate the number of people living with diagnosed HIV infection 
(PLDHIV) we performed a simple calculation using annual notifications, 
estimated mortality rates, and overseas migration rates.  

Annual HIV notifications data was provided by Australia's National HIV 
registry. Due to incomplete or inaccurate recording of name codes the 
registry contains multiple reports for some individuals. To estimate the 
number of duplicates we applied a statistical technique which has 
previously been applied to Australia's National HIV Registry . This 
calculation estimated the number of duplicate notifications annually 
resulting in 8.4% duplicate cases by 2014 with the majority of duplicates 
occurring early in the epidemic.  

We combined two approaches to estimate the number of deaths among people 
diagnosed 
with HIV infection. To estimate the number of deaths up to 2003 we used a 
linkage study conducted between Australia's National Death Index and the 
National HIV Registry 
for cases to the end of 2003 [@nakhaee2009changes]. This study calculated 
HIV- and AIDS-related
deaths and also
calculated standardized mortality ratios for people with HIV during 
different ART eras. It identified
8,519 deaths among people diagnosed with HIV or AIDS to the end of 2003. Of
these deaths, 6,900 were recorded in the HIV registry meaning 19% of all 
deaths were missing from the registry. Due to the back dating of deaths in 
the HIV registry after 2003, we used this percentage to inflated the number
of recorded deaths in the registry until the end of 2003 (inflating the 
7,102 deaths recorded to the end of 2003 to 8,768 deaths overall) and 
estimated the overall average mortality rate for PLDHIV prior to 2003. 
After 2003 we used annual mortality rates from the Australian HIV 
Observational Database (AHOD) [@ahod2014report]. Over 2004-2014, similar 
annual mortality rates were estimated for the AHOD cohort regardless of 
whether people were retained, lost or returned to follow up. We used the 
annual overall mortality rate from AHOD as the best estimate and the 95% 
confidence interval as a range in our calculations for the number of 
PLDHIV. 

We also considered the impact of overseas migration. As people are not
included in the HIV registry until they have been diagnosed in Australia 
(even if they have been
diagnosed previously overseas) we did not consider the entry of people 
living with diagnosed HIV. We
estimated an overseas migration rate for PLDHIV using data from the 
Australian Bureau of Statistics (ABS
). Given there is likely to be a flux of people leaving temporarily and 
returning to Australia (some of
which may still receive care and treatment while overseas), we used data on
the annual number of people
in the overall population who permanently leave Australia (provided by the 
ABS since 1976 in series 340102)
and the estimated resident population (ABS series 310104) to calculate an 
overall migration rate. Since 1981 this rate has risen from around 0.1% to 
0.4% of the resident population leaving Australia permanently. Due to the 
requirement for ongoing care and treatment (which is not subsidised in many
countries) we assumed a range in the annual overseas migration rate between
zero and the overall rate of permanent departure with a best estimate in 
the middle. 

Our overall estimate of the number of PLDHIV in Australia each year is 
obtained by adding the number of unique notifications to the previous 
year's estimate and subtracting the number of deaths and permanent overseas
migrants using the mortality and migration rates. 

**State and territory and sub-population estimates**

We also provided HIV estimates for the number of PLHIV and PLDHIV for each 
state and territory, mode of
exposure, region of birth, and Aboriginal and Torres Strait Islander 
status. 

For these sub-population calculations we assumed the proportion of 
duplicates, overseas migration
rate, and HIV mortality rate for each population equals the values for the 
overall population.
Mortality rates were adjusted for the Indigenous and non-Indigenous 
Australian born population to
reflect the higher overall mortality in Aboriginal and Torres Strait 
Islanders as reported by the ABS 
(http://www.abs.gov.au/ausstats/abs@.nsf/mf/3302.0).

To produce HIV cascades for each state we also considered population 
movement between states. The ABS
provides estimates for interstate arrivals and departures (series 3101016).
Using this data and estimated resident population in each state, we 
estimated the rate people leave each state and territory and the rate 
people enter from the other jurisdictions. We assumed PLDHIV move between 
states
at the same rate as the overall population as the provision of care and 
treatment is maintained across
jurisdictions. 

**Estimating the number of people living with HIV**

To estimate the overall number of MSM living with HIV (PLHIV), both 
diagnosed and undiagnosed,
we estimated the the proportion of MSM and non-MSM PLHIV who are 
undiagnosed. For MSM we used empirical
data from the COUNT study conducted alongside routine behavioural 
surveillance surveys in which gay and
homosexually active men from Sydney, Melbourne, Canberra and Perth 
recruited from a range of gay
community sites in 2013 - 2014. In this study 8.9% of participants had 
previously undiganosed HIV (95%
CI 5.8-13.5%). For non-MSM we used results from a statistical back 
projection method  which reported
20% of non-MSM living with HIV are undiagnosed [@wand2010increasing, 
@mallitt2012incidence]. For non-MSM we assumed a range of 15 to 25%.  

Multiplying the proportion of MSM and non-MSM undiagnosed by the proportion

of all diagnoses attributed
to male homosexual contact and other exposure gives a weighted average for 
the overall population of
PLHIV who are undiagnosed of 12.3%. We assumed a range of 9% to 17%. The 
overall prevalence of 
HIV in Australia was then estimated by inflating the calculated number of 
people living with diagnosed
infection by the estimated level of undiagnosed infection.

**State and territory and sub-population estimates**

We applied the national estimates for the proportion of MSM and non-MSM 
undiagnosed with HIV to each state and territory and sub-population--- 
with the non-MSM estimates were used for all non-MSM sub-populations. 


**Estimating antiretroviral treatment coverage**

We estimated the number of people receiving ART using a 10% sample of 
Pharmaceutical Benefits Scheme (PBS) patient level script claims data 
provided by the company Prospection. This is a randomised patient level, 
de-identified PBS script claims data set from 2006-present. Currently the 
data set has 170 million script claims and 3 million patients. It includes 
all PBS listed drugs with HIV 
indications. Our estimate is the number of unique patients in the PBS data 
set who filled in at least one script in the 12 months prior to the end of 
December 2014 multiplied by 10. We assumed that 10% of the Australian 
population were sampled to estimate the uncertainty range (which equates to
approximately $\pm$ 5%)

**Estimating levels of virological suppression**

We define virological suppression as less than 400 viral copies per ml. The
proportion of people
on ART with viral suppression is taken to be the proportion of people 
recorded in the Australian
HIV Observational Database (AHOD) who had less than 400 copies per ml at 
their last viral load
test.Uncertainty bounds were taken to be the proportion of people recorded 
in AHOD who had less than 1000 copies per ml and 50 copies per ml at their 
last viral load test. We estimate the number of PLHIV on ART with viral 
suppression by multiplying this proportion and range by estimated the 
number of people receiving ART. 


